---
title: "R Notebook"
output: html_notebook
---

```{r}

# Mac
#source("~/omni_credentials.R")

# PC
source("C:/Users/Malcolm/omni_credentials.R")
```


```{r}
#source("C:\Users\Malcolm\Documents\Code\GitHub")
```


#3 Connecting R to a database
##3.1 Connecting
```{r}
library(RPostgres)
```

```{r}

# source credential file
# this makes the variables omni_username and omni_password
# available to use

source("C:/Users/Malcolm/omni_credentials.R")

# create connection object using omni_username and omni_password
db_connection = dbConnect(
  drv = Postgres(), 
  user = omni_username,
  password = omni_password,
  dbname = "omni_pool",
  host = "db-postgresql-lon1-20135-do-user-7581050-0.b.db.ondigitalocean.com",
  port = 25061,
  bigint = "numeric"
)

# now we have our connection, don't need credentials anymore
# remove them from global environment
rm(omni_username, omni_password)
```

##3.2 Reading

```{r}
# list the tables, passing in the connection object
dbListTables(conn = db_connection)

# conn means connection
```


```{r}
# we can also see the fields in a table
dbListFields(conn = db_connection,
             name = "employees")
```

```{r}
# let's get the records!
data <- dbGetQuery(conn = db_connection,
                   statement = "SELECT * FROM employees")
```


```{r}
class(data)
```

Fetching records in 'chunks' - useful for big data

```{r}
result_set <- dbSendQuery(conn = db_connection,
                          statement = "SELECT * FROM employees")
class(result_set)
```

```{r}
first_10 <- dbFetch(res = result_set,
                    n = 10)
first_10
```

```{r}
next_10 <- dbFetch(res = result_set,
                    n = 10)
next_10
```

each time where cutting the data so each time we run will be the next 10

```{r}
all_remaining <-  dbFetch(res = result_set,
                          n = -1)

nrow(all_remaining)
```
Clearing the data
```{r}
dbClearResult(res = result_set)


```

##3.3 Using native SQL code chunks in R notebooks

Can use SQL code in the .Rmd, need to change the start of the code chuck

```{sql, connection = db_connection}
SELECT *
FROM employees
LIMIT 10
```


Can make the data into a variable, that gose into the Global Environment
```{sql, connection = db_connection, output.var = "out"}
SELECT *
FROM employees
LIMIT 10
```



```{r}
library(tidyverse)

out %>% 
  summarise(avg_salary = 
              mean(salary))
```

##3.4 Disconnecting

Finally, it is good practice to close the database connection, once we are finished with it.

```{r}
dbDisconnect(conn = db_connection)
```

