---
title: "R Notebook"
output: html_notebook
---

```{r}
# incomes <- read_csv("data/state_income_data.csv")
```

# Question 1. Load in the data, and have a look at it. How many variables does it have? How many observations? What are the variable names? What type of variables does it have?

```{r}
library(CodeClanData)
library(tidyverse)

incomes <- state_income_data
incomes
```

#Question 2. Select the columns which contain the information from years 2002 to 2008. Ensure you keep the Index and State variables as well.

```{r}
six_years <-  select(incomes, 1:9)
six_years
```

#Question 3. Rename the Index column to first_letter. Hint: look up the rename() function.

```{r}
six_years <- rename(six_years, first_letter = Index)
six_years
```

#Question 4. Find the seven states which start with an A,B, or C.

```{r}
A_to_C_states <- filter(six_years, first_letter == "A" |
                         first_letter == "B" |
                         first_letter == "C")
A_to_C_states
```

#Question 5. Find which of these seven states have the greatest income in 2008.

```{r}
arrange(A_to_C_states, desc(Y2008))
```

#Question 6. Calculate the change in income for each state of these seven states (states starting with A,B, or C) between 2002 and 2008. Save the income change in a new column called income_change.

```{r}
A_to_C_states  <- 
                 mutate(A_to_C_states, income_change = Y2008-Y2002)


A_to_C_states%>% 
                 mutate(income_change = Y2008-Y2002)
# two way to do this
```

#Question 7. Find which of these seven states income change is greater than 500,000 (i.e. an increase of 500,000 or decrease of 500,000).

```{r}
greater_change <- A_to_C_states %>% 
                  filter(income_change > 500000 |
                         income_change < -500000)
greater_change
```

# Question 8. Calculate the mean and median income change for all seven states.

```{r}
median_income_change <- A_to_C_states%>%
                        summarise(median = median(income_change))
median_income_change

mean__income_change <- A_to_C_states%>%
                        summarise(mean = mean(income_change))

mean__income_change
```

#Question 9. Go back to the original dataset which contains all states. Write a pipe which selects the State and Y2008 variables, arranges the income from highest to lowest, filters all those states which have a income of greater than 1 million, and then finally chooses the top 10 wealthiest states.

```{r}
Top_states <- incomes %>% 
              select(State, Y2008) %>% 
              arrange(desc(Y2008)) %>% 
              filter(Y2008 > 1000000) %>% 
              head(10)
Top_states
```

