---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```
 
```{r}
single_string <- "string a"
strings <- c("string a", "string b", "string c", "string d", "striiiiiiing e")

# all start with str
# all have ing in it
# all end with a single letter
```

```{r}
patten <- "a"
```


```{r}
single_string %>% 
  str_detect(patten)
```

```{r}
strings %>% 
  str_detect(patten)
```

```{r}
pattern <- "str"
strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "[abc]" # in the [] dose it have "a" or "b" or "c"
strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "[a-c]" # looking for letters between a to c
strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "[A-Z]" # looking for CAPS letters between A to Z, which we have none
strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "[a-zA-Z]" # looking for CAPS and lower letters between A to Z
strings %>% 
  str_detect(pattern)
```

```{r}
strings
```

```{r}
pattern <- "i{3}" # find 3 consecutive letter i's (lowcase)
strings %>% 
  str_detect(pattern)


pattern <- "i{3}" 
strings %>% 
  str_detect("i{3}")  # if the pattern is small you can use in the code
                      # best to put the pattern at the top 



```


# Task
Make a vector containing the following strings:
“We’re having great fun learning about data analysis at CodeClan!”
“CodeClan is at 37 Castle Terrace in Edinburgh, ClockWise in Glasgow, and Wasps Academy in Inverness”
“Our cohort code starts with DE, DG or D”
“You get a great view of the castle or the national conservatoire from just outside CodeClan’s door”
Write and test regular expressions which will match:
Any string mentioning CodeClan
Any string containing two consecutive capital letters
Any string not including the word “castle” or “Castle”
12:04
Hint
 str_detect() can take an optional third argument called negate
12:05
Hint
The techniques we looked at above can be combined.



```{r}
test_strings <- c("We’re having great fun learning about 
                  data analysis at CodeClan!",
                  "CodeClan is at 37 Castle Terrace in 
                  Edinburgh, ClockWise in Glasgow,
                  and Wasps Academy in Inverness",
                  "Our cohort code starts with DE, DG or D",
                  "You get a great view of the castle or the 
                  national conservatoire from just outside 
                  CodeClan’s door")
```

```{r}
codeclan <- "CodeClan"    # Finds the exact
test_strings %>%
  str_detect(codeclan)


caps <- "[A-Z]{2}"        # Finds 2 consecutive letter in UPPER CASE 
test_strings %>%
  str_detect(caps)

no_castle <- "[cC]astle"  # Finds castle with a UPPER or lower case
test_strings %>%
  str_detect(no_castle, negate = TRUE)  # Finds strings not including castle


  !str_detect(test_strings, "[cC]astle") # Same as above, from Ricardo

```


## Extracting matching substrings


[] - or operation
{2} -  number of occurrence
(?i) - [un]capalisation finds if it upper or lower case
* - zero or more occurrence
+ - one or more occurrence


```{r}
strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)
```

```{r}
test <- "THIS is a string"
pattern <- "(?i)[a-z]"   #"[a-zA-Z]" is the same "(?i)[a-z]" Finds anything with a UPPER or lower case

test %>% 
  str_extract(pattern)
```


```{r}
pattern <- "[a-z]"
strings %>% 
  str_extract(pattern)

# This if finding the first lower case letter
```


```{r}
strings

pattern <- "[a-z]@"
strings %>% 
  str_extract(pattern)
```

```{r}
invalid_email <- "@emailproider.com"
pattern <- "[a-z]*@emailproider.com"

invalid_email %>% 
  str_extract(pattern)

# This will find everything before the @ including nothing 
```

```{r}
invalid_email <-  "@emailprovider.com"
pattern <- "[a-z]+@emailprovider.com"

invalid_email %>% 
  str_extract(pattern)

# This will only find if there's something before the @ 
```


```{r}
strings

pattern <- "[a-z]+@"
strings %>% 
  str_extract(pattern)
```


```{r}
pattern <- "[a-z0-9]+@[a-z]+"
strings %>% 
  str_extract(pattern)
```


```{r}
pattern <- "[a-z0-9_\\.]+@[a-z0-9_\\.]+" # This finds all letters, numbers, underscore and full stop
strings %>% 
  str_extract(pattern)

# \\. mean ignore . meaning everything just read it as a full stop
```

```{r}
pattern <- "[a-z0-9_\\.]+@[a-z0-9_\\.]+" 
strings %>% 
  str_extract_all(pattern)
# To find the 4 email address use this
# This makes a matrix will have a lesson on this later
```

```{r}
pattern <- "[a-z0-9_\\.]+@[a-z0-9_\\.]+" 

strings %>% 
  str_replace(pattern, "REDACTED")

# This replaces the email with "REDACTED" but only the first in the each line
```

```{r}
pattern <- "[a-z0-9_\\.]+@[a-z0-9_\\.]+" 

strings %>% 
  str_replace_all(pattern, "REDACTED")

# This replaces the email with "REDACTED"
```

Task - 15 mins 
Make a vector containing the strings:
“Colin’s mobile number definitely isn’t 07777777777.”
“A landline might be 01010 123456.”
“Mhairi, Del, Aileen and Steph have been some of your instructors so far.”
“I’m called your_name_here and I live in your_town_here.”
Use regular expressions to:
Extract all the names
Extract the landline number
Insert your name and where you live into the fourth string

```{r}
lloyd <- c("Colin’s mobile number definitely isn’t 07777777777.",
           "A landline might be 01010 123456.",
           "Mhairi, Del, Aileen and Steph have been some of your instructors so far.",
           "I’m called your_name_here and I live in your_town_here."
           )
```

```{r}
# extract_name <- "[A-Z]+a-z"
# lloyd %>% 
#   str_extract(extract_name)

extract_name <- "[A-Z][a-z]+"         # Upper case followed by all lower case 

extracted <- lloyd %>%
  str_extract_all(extract_name)

extracted %>% 
  flatten_chr()     # trues it from a matrix into a var again
```


```{r}
extract_number <- "[0-9]+|[0-9]+ [0-9]+"   # better to wright as"[0-9]+ [0-9]+|[0-9]+"
lloyd %>%                                  # as it's including the space before " 01010
  str_extract(extract_number)

extract_number <- "[0-9]{5} [0-9]{6}|[0-9]{11}"
lloyd %>% 
  str_extract(extract_number)




#"[0-9]{5} [0-9]{6}|[0-9]{11}"


```


```{r}
my_address <- "Dunblane"
my_name <- "Mal"
lloyd %>% 
  str_replace("your_town_here", my_address) %>% 
  str_replace("your_name_here", my_name)
```


# Handy Regex Notations

?!':;,

```{r}
punct_strings <- c(
  "Why is this a question?", 
  "I am a sentence with no punctation",
  "I am a well punctuated sentence.", 
  "WOW!"
)
```


```{r}
punct_strings %>% 
  str_extract_all("[:punct:]")
```


```{r}
punct_strings %>% 
  str_extract_all("[A-Z]{3}[:punct:]") %>% 
  flatten_chr()
```

^ - start of the string
$ -end of the string


```{r}
test_strings <- c(
  "abc", 
  "Abc",
  "abC",
  "s"
)
```


```{r}
test_strings %>% 
  str_extract_all("[a-z]")
# looking for words that have a lower case
```

```{r}
test_strings %>% 
  str_extract_all("^[a-z]")
# looking for words that start with a lower case
```

```{r}
test_strings %>% 
  str_extract_all("[a-z]$")
# looking for words that end with a lower case
```

```{r}
test_strings %>% 
  str_extract_all("^[a-z]$")
# Finds words the start and finish with a lower case
```

```{r}
test_strings %>% 
  str_extract_all("^[a-z]+$")

# The one above is just looking for words with just 1 chr [a-z], adding the [a-z]+ looiking for any word 
```

```{r}
student <- c("Abigail", "Bjorn", "Caitlin")
age <- c(25, 56, 40)
emails <- c("abigal12@company.co.uk", "bsmith@gmail.co.uk", "caitlin_1980")

student_info <- tibble(student, age, emails)

student_info 
```

```{r}
student_info %>% 
  mutate(email_detect = str_detect(emails, "[a-z0-9]+@"))
```

```{r}
student_info %>% 
  mutate(emails = str_extract(emails, "[a-z0-9_\\.]+@[a-z\\.]+"))
```


