---
title: "Using dates and times"
output: html_notebook
---

- formatting them in programming can be tricky
- lots of different ways to enter them
- tricky data format to work with

ISO 8601 format



### Task - 5 mins 
Take a few minutes and read about IS0 8601 standard, as we will be working with it a lot today. Answer the following questions:

    What order does ISO standard follow? (e.g. year to second, second to year…)
    What is the minimum number of digits you can have in a year? Why is this?
As a minimum, a four-digit year [YYYY] to avoid the year 2000 problem.
    
    What clock system do the times use?
uses the 24-hour clock system.
    
    How many time formats does ISO standard have?
6
    
    What timezone are time and dates automatically represented in?
UTC universal coodinated time GMT
    
    How are combined date-times represented?
2021-12-13T10:13



```{r}
library(lubridate)
```

```{r}
bday <- "09-01-2021" 

class(bday)
```

```{r}
bday_change <- dmy(bday)

class(bday_change)
```

```{r}
mdy("January 9th, 2021")
```


```{r}
library(tidyverse)

netflix <- read_csv("data/netflix.csv")
```

```{r}
glimpse(netflix)
```

### Task - 5 mins

Change the column date_added from character to a date, using either dmy, mdy or ymd.

```{r}
netflix <- netflix %>% 
  mutate(date_added = mdy(date_added))

netflix
```

```{r}
netflix <- netflix %>% 
  mutate(day = day(date_added),
         month = month(date_added, 
                       label = TRUE, abbr = FALSE),
         year = year(date_added)
  )

netflix

```


### Task - 5 mins

Add a column weekday, which is the day of the week the title was added. We want this as the abbreviated day of the week i.e. Mon, Tue etc.

```{r}
netflix %>% 
  mutate(weekday = weekdays(date_added, 
                             abbr = TRUE))
```


```{r}
netflix <- netflix %>% 
  mutate(made_date = make_datetime(year, month, day))

head(netflix)
```

### Task - 5 mins

Look up how the make_datetime function works. Redo your column, but add a time onto it as well. You can pick what time you add - try out a few combinations to see how it’s working. Don’t worry about time zones for now.
```{r}
netflix %>% 
  mutate(time = make_datetime(year, month, day, tz = "UTC"))
```
johny work
```{r}
netflix <- netflix %>%
  mutate(made_date = make_datetime(year, month, day, hour = 12, min = 12, sec = 15, tz = "EST" ))

head(netflix) %>% 
  select(made_date) %>% # just shows the date and time
  pull() # shows the "EST"
```

# 9 Calculating time spans

- Difftime: a datetime object that represents days
- Period: represent a timespan like weeks, months and days
- Duration: represent an exact number of seconds


Calculate how many days a movie/tv show has been released for
```{r}
netflix %>% 
  mutate(time_diff = today() - date_added) %>% 
  select(time_diff, date_added) %>% 
  head()
```

Periods :
```{r}
netflix %>% 
  mutate(time_period = as.period(today() - date_added)) %>% 
  select(time_period, date_added) %>% 
  head()
```

```{r}
netflix %>% 
  mutate(time_period = as.period(today() - date_added),
         time_years = as.numeric(time_period, "years"),
         time_hours = as.numeric(time_period, "hours")
  ) %>% 
  select(time_period, time_years, time_hours, date_added) %>% 
  head()
```


