---
title: "Distributions - measures of spread"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(skimr)
```

```{r}
jobs <- read_csv("data/TyrellCorpJobs.csv") %>% 
  clean_names()
jobs
```

# Measures of spread

## Range

range = max_value - min_value

spread=max. value_min. - value


range of salaries is just the max salary - min salary
```{r}
jobs %>% 
  summarise(max = max(salary),
            min = min(salary),
            range = max(salary) - min(salary))

# n.b. R has a range function - it returns max and min but dosn't coulate the diff
```

```{r}
range(jobs$salary)
```

```{r}
diff(range(jobs$salary))
```


```{r}
jobs %>% 
  ggplot() +
  geom_histogram(aes(x = salary),
                 col = "white", bins = 25)
```

```{r}
# separating position
jobs %>% 
  ggplot() +
  geom_histogram(aes(x = salary),
                 col = "white", bins = 25) +
  facet_wrap(~position)
```

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(range = diff(range(salary)))
```

A potential problem (with using the range as a measure of spread): it is very strongly affected by outliers.

## Quartiles and the Interquartile Range

Q1: the value splitting the distribution such
that 25% data below, 75% above
Q2: "" "" "" such that 50% data below 50% above - the median
Q3: "" "" "" such that 75% data below 25% above
--|--|--|--
```{r}
# quartiles for the jobs data
jobs %>% 
  group_by(position) %>% 
  summarise(
    Q1 = quantile(salary, 0.25),
    Q2 = quantile(salary, 0.5),
    Q3 = quantile(salary, 0.75)
  )
```
```{r}
iqr <- jobs %>% 
  group_by(position) %>% 
  summarise(
    Q1 = quantile(salary, 0.25),
    Q2 = quantile(salary, 0.5),
    Q3 = quantile(salary, 0.75)
  ) %>% 
  pivot_longer(Q1:Q3, names_to = "fencepost", values_to = "value")
jobs %>% 
  ggplot() +
  geom_histogram(aes(x = salary), col = "white", bins = 25) +
  geom_vline(xintercept = iqr$value, linetype = "dashed", colour = "red") +
  facet_wrap(~position)
```
## The Interquartile Range
The range between the fenceposts Q3 and Q1
```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(
    Q1 = quantile(salary, 0.25),
    Q2 = quantile(salary, 0.5),
    Q3 = quantile(salary, 0.75),
    IQR = Q3 - Q1,
    # R also has an IQR function:
    IQR_fun = IQR(salary)
  )
```

# The five figure summary

min, Q1, Q2, Q3, max
```{r}
summary(jobs$salary)
```
We wanted accounting and management to be grouped
though
```{r}
library(skimr)
```
`skim()` is a more sophisticated summary
```{r}
ffs <- jobs %>% 
  select(-x1) %>% 
  group_by(position) %>% 
  skim()
```
```{r}
ffs$numeric.hist
```

## Box Plots

An excellent statistical diagnostic tool
box plots / hinge plots / box and whisker / tukey 
```{r}
jobs %>% 
  ggplot() +
  aes(x = salary, y = position) +
  stat_boxplot(geom = "errorbar", width = 0.5) +
  geom_boxplot()
```
- a more rigourous definition of what an outlier is
whiskers are typically positioned 1.5 times the IQR above and below Q3 and Q1 respectively. Any points that lie outside of that range are outliers

## Skew in box plots:

What does skew look like on a box plot?
```{r}
heavily_right_skewed <- read_csv("data/heavily_right_skewed.csv")
```
```{r}
heavily_right_skewed %>% 
  ggplot() +
  geom_boxplot(aes(x = x))
```
```{r}
heavily_right_skewed %>% 
  ggplot() +
  geom_histogram(aes(x=x), colour = "white", bins = 30)
```
```{r}
library(e1071)
```
```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(skewness = skewness(salary, type = 1))
```
break - 5 mins
Aside
- remove outliers from the data (potentially dangerous)
- get where Q1 and Q3 are
- get interquartile range
- calculate distance to whisker (1.5*IQR)
- filter values above Q3 + 1.5IQR, values below Q1 - 1.5IQR

# Variance

- a single measure of spread
variance = how far each value is from the mean
variance of a sample = $s^2$
sHow to calc variance step by step:
1. get the difference between each value and the mean - squared
2. sum them
3. divide by the total number of values - 1
How to calc variance in R:
`var()`
```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(variance = var(salary))
```
dollars squared doesn't mean a lot to me. How many pounds is that worth?

## Standard Deviation

- a more interpretable single measure of spread of the data
to go from variance to st. dev all we do is take
the square root
```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(mean = mean(salary),
            # david proving R works as expected
            st_dev = sqrt(var(salary)),
            sd = sd(salary)) 
```
so... if you come and work for TyrellCorp in
Accounting, you can expect to earn 34,115 +/- 2400
Management, you can expect to earn 51,000 +/- 11000
