---
title: "Distributions"
output: html_notebook
---

```{r}
library(prob)
library(tidyverse)
```

# Distributions

### A rigourous definition

A distribution is a list of all of the possible outcomes of a random variable along with either their corresponding frequency or probability values (we would speak of the frequency distribution or probability distribution respectively).

### Less rigourous

Defined either way, the distribution gives insight into how likely or how common the various outcomes are.

```{r}
three_coins <- tosscoin(3)

three_coins
```

#### Task - 2 mins what is the probability of each of the eight outcomes if all of the coins are unbiased? What does ‘unbiased’ mean here? Why is it important that each of the coins are unbiased in this calculation?


Distribution 1: Uniform Distribution
```{r}
three_coins_p <-  three_coins %>% 
  mutate(label = str_c(toss1, toss2, toss3)) %>% 
  mutate(prob = 1/n())

three_coins_p %>% 
  ggplot() +
  geom_col(aes(x = label, y = prob))
```

Uniform Distribution - 

All the bars are the same height, indicating the probability of each outcome is the same

A bit more intrtrsting ...

What are the probabilities of getting 0 1, 2 or 3 heads on the toss of three unbiased coins? 

```{r}
three_coins_p %>% 
  mutate(n_heads = str_count(label, "H")) %>% 
  group_by(n_heads) %>% 
  summarise(prob = sum(prob))
```

```{r}
three_coins_p %>% 
  mutate(n_heads = str_count(label, "H")) %>% 
  group_by(n_heads) %>% 
  summarise(prob = sum(prob)) %>% 
  ggplot() +
  geom_col(aes(x = n_heads, y = prob)) + 
  xlab("Number of heads thrown") + 
  ylab("\n Probabiltiy \n")
```

Distribution is no longer uniform (bars havediffering hights)

Distribution is still discrete (bar-able)

General Properties of Discrete Distributions

x-axis = outcomes
y-axis = probaility or frequency
THE PROBABILITIES MUST SUM TO 1

If you stacked all the bars on top of each other, the y-axis would read 1.


## Measures of centrality: mean, median and mode

```{r}
library(janitor)
library(lubridate) # for date and times wrangling
```

```{r}
air_con_sales <- read_csv("data/AirConSales.csv") %>% 
  clean_names()

air_con_sales
```

```{r}
air_con_sales <- air_con_sales %>% 
  mutate(date = mdy(date)) # this is the oder the data is in mdy

air_con_sales
```

We've got mesurements for each day, but we want to look at how likely it will be to sell a certain number of units (air conditioners)

Generating a frequency table

```{r}
sales_freq_table <- air_con_sales %>% 
  tabyl(units_sold)

sales_freq_table
```


```{r}
sales_freq_table %>% 
  ggplot() +
  geom_col(aes(x = units_sold, y = n))
```

`mean` - sum all of values / n of all values

`median` - 50% of data beneath, 50% of data above

`mode` - most frequent value

R has a mean and a median function, but not a mode function


```{r}
get_mode <- function(data){

  tabled_data <- table(data)
  table_names <- names(tabled_data)
  
  return(as.numeric(table_names[tabled_data == max(tabled_data) ]))
  
}
```

Let's calculate these values for our distribution

```{r}
air_con_sales %>% 
  summarise(mean_daily_sales = mean(units_sold),
            median_daily_sales = median(units_sold),
            mode_daily = get_mode(units_sold))
```

Bill Gates effect: the mean is more heavily skewed by outliers

Extremes inform the means - Jonny 2021 :p

Measures of centre and outliers

If you suspect or observe outliers in your data, be carful which measure of centre you use. Always look at the distribution and distribution 

unimodal just means ‘one-peaked’

bimodal means ‘two-peaked’


Skeqness - assymetry of a distrbution

left-skew = left-tailed

the mean is less than median

right-skew = right-tailed

the mean is greater than the median

skew indicates where the mean is relative to the median


Calculating skewness
```{r}
╔library(e1071)
```

```{r}
left_skewed <- read_csv("data/leftskew.csv")
right_skewed <- read_csv("data/rightskew.csv")

left_skewed %>% 
  summarise(skewness = skewness(x, type = 1))

right_skewed %>% 
  summarise(skewness = skewness(x, type = 1))
```
left_skew = -1
right_skew = ~ 0.7

direction - positive/negative

negtive skew = left skew

magnitude - 
0 - 0.5 = fairly symmetrical
0.5 - 1 = moderately skewed
1 or greater = highly skewed



```{r}
air_con_sales %>% 
  summarise(skewness = skewness(units_sold, type = 1))

# The distribution is highly right-skewed. 
```

The Key Takeaway

*It's important to visualise your data**

We’ve been saying this a lot, but it bears repeating:

“Always visualise your data before performing any further statistical analyses!” 

