---
title: "R Notebook"
output: html_notebook
---

# Logistic Regression

binary dependent variable: (0 or 1)

- spam emails or not
- married or not
- clicked or not
- cat or dog
- organic or conventional 
- registered user or not
- accepted or not mortgages


```{r}
library(tidyverse)
library(janitor)
```

```{r}
mortgage_data <- read_csv("data/mortgage_applications.csv") %>%
  clean_names()
              
head(mortgage_data)
```

```{r}
library(GGally)

ggpairs(mortgage_data)
```

```{r}
# Let's reduce the symbol size and 'jitter' the y-values so we can see more of the data without overlap of symbols
# geom_jitter() adds adds a small amount of random variation (vertically and/ore horizontally depending on the arguments) to the location of each point

score_plot <- ggplot(mortgage_data) +
  geom_jitter(aes(y = tu_score, x = as.integer(accepted)), shape = 1, 
              position = position_jitter(width = 0.05)) + 
  xlab("Accepted Status") + scale_x_continuous(breaks=seq(0, 1,1))

score_plot
```

# Logistic Function

Logistic regession is estimating the LOG ODDS OF AN EVENT, RATHER THAN THE PROBABILTITY.

ask - 2 mins
If we roll a fair die, what are the odds of getting a 6?

Odds(success) = 1/6 / 5/6 = 0.2

```{r}
logit <- function(x){
  return(log(x/(1-x))) # x is prob
}

logit_data <- tibble(p = seq(0.001, 0.999, 0.001)) %>%
  mutate(logit_p = logit(p))

head(logit_data)
```

```{r}
ggplot(logit_data, aes(y = p, x = logit_p)) + 
  geom_line() + 
  ylab("probability") + xlab("logit(p) value")
```


**Build a model that predicts the probaility of a loan being accepted or not, given a tu_score**

1. Build the logistis model

```{r}
mortgage_data_log_model <- glm(accepted ~ tu_score, data = mortgage_data, 
                               family = binomial(link = "logit"))

mortgage_data_log_model
```

2. Add these predictions into your dataset

```{r}
library(modelr)

predict_log <- tibble(tu_score = seq(0, 710, 1)) %>% 
  add_predictions(mortgage_data_log_model, type = "response")
```

3. Check model fits the data?
```{r}
ggplot(mortgage_data) + 
  geom_jitter(aes(x = tu_score, y = as.integer(accepted)), shape = 1, 
              position = position_jitter(h = 0.05)) + 
  geom_line(data = predict_log, aes(x = tu_score, y = pred), col = 'red')
```


Task - 5 mins
Use and amend the code above to predict the probability of getting a mortgage application accepted with a tu_score of 594.

```{r}
predict_log_task <- tibble(tu_score = 594) %>% 
  add_predictions(mortgage_data_log_model, type = "response")

predict_log_task
```

60% of people will get a mortgage application accepted 

1/1 + e(b0 = b1 * x)

```{r}
ggplot(mortgage_data) + 
  geom_jitter(aes(x = tu_score, y = as.integer(accepted)), shape = 1, 
              position = position_jitter(h = 0.05)) + 
  geom_point(data = predict_log_task, aes(x = tu_score, y = pred), col = 'red')
```

4. Interpreting the B1 for a continuous predictor (tu_score)

If the independent variable increases by one unit, then the estimate of the log odds of suscces changes (increases or decreases) by B1 units.

Calculate the odds of having an accepted application at tu_score = 594:
```{r}
odds_at_594 <- tibble(tu_score = 594) %>% 
  add_predictions(mortgage_data_log_model, type = "response") %>% 
  mutate(odds = pred / (1-pred))

odds_at_594
```

This implies that a 1 unit increase in tu_score increases the odds of getting approved for a loan by a factor of 1.58.

How do these odds change if we increase the tu_score by 50 points?


```{r}
library(broom)
```

```{r}
mortgage_data_log_model
```
Take the above details and make a table to read it easier below

```{r}
b_tu_score <- tidy(mortgage_data_log_model) %>% 
  filter(term == "tu_score") %>% 
  select(estimate)

b_tu_score
```

```{r}
odds_factor <- exp(b_tu_score * 50)

odds_factor
```

Calculate the new odds of getting a loan accepted or not, based on a 50 point increase in tu_score

```{r}
odds_at_594 <- odds_at_594 %>% 
  select(odds)

odds_at_594
```

```{r}
new_odds <- odds_factor * odds_at_594

new_odds
```

```{r}
odds_at_644 <- tibble(tu_score = 644) %>% 
  add_predictions(mortgage_data_log_model, type = "response") %>% 
  mutate(odds = pred / (1-pred))

odds_at_644
```

Task - 5 mins How do the odds of acceptance change if we decrease tu_score by 50 points from 594 to 544? Calculate the new odds using R.

[Hint - the change in x here is −50] 

```{r}
odds_at_544 <- tibble(tu_score = 544) %>% 
  add_predictions(mortgage_data_log_model, type = "response") %>% 
  mutate(odds = pred / (1-pred))

odds_at_544
```


