---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
comms_data <- read_csv("data/telecom_data.csv")


# select comms_data then hit ctrl + enter to get the table 

# Chris had problems with his monthly_charges beaning a chrater than a number
# comms_data <- read_csv(“data/telecom_data.csv”,
#                        col_types = cols(“monthly_charges” = col_number()))
```

```{r}
comms_data %>% 
  filter(is.na(monthly_charges))
```

```{r}
comms_data %>% 
  summarise(count = sum(is.na(monthly_charges)))
```
# Get rid of the N/A rows, will lose data 
```{r}
comms_data %>% 
  drop_na()
```
# if there's N/A it will throw off the mean/ median
```{r}
comms_data %>% 
  mutate(monthly_charges = coalesce(
    monthly_charges,
    median(monthly_charges)
    ))

summary(comms_data)
```

# This will add values to the N/A simalar to the other values
# Median
```{r}
comms_data %>% 
  mutate(monthly_charges = coalesce(
    monthly_charges,
    median(monthly_charges,
    na.rm = TRUE)
    ))
```

# Mean
```{r}
comms_data %>% 
  mutate(monthly_charges = coalesce(
    monthly_charges,
    mean(monthly_charges,
    na.rm = TRUE)
    ))
```

# task find all the NA's
```{r}
comms_data %>% 
  filter(is.na(total_charges))
```

# reading the table I can see more NA
# total_charges is a chrater rather than number
```{r}
comms_data %>% 
  summarise(count = sum(is.na(total_charges)))
```
# Had a look and seen na and N/A, this changes them to the same NA 
```{r}
total_charges_replaced <- comms_data %>% 
  mutate(
    total_charges = na_if(total_charges, "na"),
    total_charges = na_if(total_charges, "N/A")
    )

  # total_charges = na_if(total_charges, "na|N/A") dose the same
```
# Now finds all 9 rows
```{r}
total_charges_replaced %>% 
  filter(is.na(total_charges))
```
# Changes total_charges from CHR to DBL
```{r}
total_charges_replaced %>% 
  mutate(total_charges = as.numeric(total_charges))

```

# same as total_charges, seen NA & -- used, changed to NA
```{r}
payment_method_replaced <- total_charges_replaced %>% 
  mutate(
    payment_method = na_if(payment_method, "NA"),
    payment_method = na_if(payment_method, "--")
    )
```


```{r}
payment_method_replaced %>% 
  filter(is.na(payment_method))
```
# changes the NA to "unavailable"
```{r}
payment_new_category <- payment_method_replaced %>% 
  mutate(payment_method = coalesce(payment_method, "unavailable"))
payment_new_category
```


