---
title: "Over-fitting"
output: html_notebook
---


```{r}
library(tidyverse)
library(CodeClanData)
savings
```

# Over-fitting in multiple linear regression

```{r}
model_overfit <- lm(savings ~ ., data = savings)

model_wellfit <- lm(savings ~ salary + age + retired, data = savings)

model_underfit <- lm(savings ~ salary, data = savings)
```


```{r}
summary(model_overfit)

summary(model_wellfit)

summary(model_underfit)
```

# Parsimonious measures of goodness-of-fit

A Parsimonious model is one which included as few variables as necessary.

Adjusted R squared - Measures the proportion of the variation in your response variable

Akaike Information Criterion (AIC) - Single number score used to determine which of the model is most likely the best model.

Bayes Information Criterion (BIC) - Measures of goodness of fit

R-squared - Larger values are better
AIC and BIC - Lower numbers are better

BIC tends to be more parsimonious. (It tends to select smaller models)
 - BIC penalizes larger models when compared to AIC



```{r}
summary(model_overfit)$adj.r.squared
```

```{r}
AIC(model_overfit)
```

```{r}
BIC(model_overfit)
```

If you want to see all the scores in one go in a handy data-frame form, you can install the broom package and use the glance function.

```{r}
broom::glance(model_overfit)
```

Task - 10 minutes

Find the R-squared, adjusted R-squared, AIC and BIC score for:

    The well fitted model

model_wellfit <- lm(savings ~ salary + age + retired, data = savings)

```{r}
summary(model_wellfit)$adj.r.squared

AIC(model_wellfit)

BIC(model_wellfit)
```

    The over fitted model

model_overfit <- lm(savings ~ ., data = savings)

```{r}
summary(model_overfit)$adj.r.squared

AIC(model_overfit)

BIC(model_overfit)
```

    The under-fitted model

model_underfit <- lm(savings ~ salary, data = savings)

```{r}
summary(model_underfit)$adj.r.squared

AIC(model_underfit)

BIC(model_underfit)
```


Does the results you found match with your expectations?

model_wellfit has the lower AIC & BIC and the mid r^2 which will be the better one 


jonny's work
```{r}
a <- broom::glance(model_overfit)

b <- broom::glance(model_underfit)

c <- broom::glance(model_wellfit)

models_summary_comparison <- bind_rows(a,b,c)

models_summary_comparison
```

## Test and traing set


Exploratory Data Analysis - EDA

Feature Engineering

Feature Selection 

Frature Store

Test and traing set

Model Development






Below is an example of splitting a dataset into a test and train set. First you use random sampling to get the indexes of the test data, then you can use slice to split the data. Let’s do this on our savings data.

```{r}
set.seed(9)

# count rows
n_data <- nrow(savings)

# make teat index
test_index <-  sample(1:n_data, size = n_data*0.2)

# use test index to create test and training split
test <- slice(savings, test_index)
train <- slice(savings, -test_index)

```

Now we can fit the model on the the train data:

```{r}
model <- lm(savings ~ salary + age + retired,
            data = train)
```

Using that model we can make predictions based on our test data. We can use the add_predictions() function from modelr to add a prediction for every observation in our dataset.

```{r}
library(modelr)

predictions_test <- test %>%
  add_predictions(model) %>%
  select(savings, pred)

predictions_test
```

Now we can calculate the mean squared error, just by taking the the average of the squares of the differences between the predictions and the actual values.

```{r}
mse_test <- mean((predictions_test$pred - test$savings)**2)
mse_test

# can use either vaible

# mse_test <- mean((predictions_test$pred - predictions_test$savings)**2)
# mse_test
```


Task - 10 minutes

Calculate the mean squared error between predicted savings and actual savings in the training dataset.

Which is higher, then error on the test or the error on the training data? Is this what you would expect?

```{r}
predictions_train <- train %>%
  add_predictions(model) %>%
  select(savings, pred)

predictions_train
```

```{r}
mse_train <- mean((predictions_train$pred - predictions_train$savings)**2)
mse_train 
```

## K-fold cross validation

```{r}
library(caret)
```

```{r}
# set options for cv
cv_10_fold <- trainControl(method = "cv",
                           number = 10,
                           savePredictions = TRUE)

model <- train(savings ~ salary + age + retired,
               data = savings,
               trControl = cv_10_fold,
               method = "lm")
```

```{r}
model$pred
```

```{r}
model$resample
```

```{r}
mean(model$resample$RMSE)
```

```{r}
mean(model$resample$Rsquared)
```





Ricardo's work
```{r, message=FALSE, warning=FALSE}
cv_10_fold <- trainControl(method = "cv",
                           number = 10,
                           savePredictions = T)

model <- train(savings ~ .,
               data = savings,
               trControl = cv_10_fold,
               method = "lm")

mean(model$resample$RMSE)
```

```{r, message=FALSE, warning=FALSE}
cv_10_fold <- trainControl(method = "cv",
                           number = 10,
                           savePredictions = T)

model <- train(savings ~ .,
               data = savings,
               trControl = cv_10_fold,
               method = "lm")

mean(model$resample$Rsquared)
```


rout mean squred error - RMSE

well-fit model RMSE - 9663.584

over-fit model RMSE - 9789.31



well-fit model Rsquared - 0.340796

over-fit model Rsquared - 0.3216934


Test-train split is 20% - 80%

K-fold cross validation uses the 80% to make x number of folds (ie 5 or 10 samples from the 80% of the data, 3 should be the minimum), the mean of that is taken to average out the errors 

In the examples above we're using the whole data to test the model but would normally uses the 20% validation set that we split off. See the next part



```{r}

```





