---
title: "Clustering Lab"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(CodeClanData)
```



Part a

* We’ve been asked to create clusters of students for the answers to the ‘importance’ questions (where students rate how important they find certain topics on a scale from 0 to 1000, the higher the number the more important they think the issue is).

* If we are to use the mathematical methods (elbow etc.) what would your optimal value of k been? Do you think the data seems well suited to cluster.

```{r}
students_big
```

```{r}
glimpse(students_big)
```


```{r}
scale_students <- students_big %>% 
                      select(15:20) %>% 
                      # pivot_longer(
                      #   cols = starts_with("importance_"),
                      #   names_to = "question",
                      #   names_prefix = "importance_",
                      #   values_to = "importance",
                      #   values_drop_na = TRUE) %>% 
                      mutate(across(where(is.numeric), scale))# %>% 
                      column_to_rownames("question") 

scale_students
```

```{r}
set.seed(1234)

clustered_edu <- kmeans(scale_students,
                        centers = 6,
                        nstart = 25)

clustered_edu
```

```{r}
augment(clustered_edu, students_big %>% 
                      select(15:20))
```

```{r}
# augment(clustered_edu, students_big %>% 
#                       select(15:20)) %>% 
# pivot_longer(
#   cols = starts_with("importance_"),
#   names_to = "question",
#   names_prefix = "importance_",
#   values_to = "importance",
#   values_drop_na = TRUE) 
```

```{r}

```

