---
title: "Hierarchical Clustering"
output: html_notebook
---

```{r}
library(tidyverse)
library(cluster)
library(factoextra)
library(dendextend)
library(corrplot)
```

```{r}
edu_data <- read_csv("data/school_data.csv")
```

```{r}
head(edu_data)
```

```{r}
# check the data. For clustering, we want no text, but want to be able to view the states still. We can get round this, by moving the column (X1) which has our text, into the rownames instead. 
edu_data <- edu_data %>%
  column_to_rownames("X1")
head(edu_data)
```

So we have a data frame, with 50 rows and 5 columns. Next we can check if we have any missing values to deal with.
```{r}
any(is.na(edu_data))
```

```{r}
edu_data_scale <- edu_data %>%
              mutate_if(is.numeric, scale)

edu_data_scale <- edu_data %>%   # dose the same but is more up todate
              mutate(across(where(is.numeric), scale))

edu_data_scale
```

```{r}
edu_data_scale %>%
  pivot_longer(cols = everything(),
               names_to = "type", 
               values_to = "value") %>% #convert data to long format
  group_by(type) %>%
  summarise(mean = round(mean(value)), 
            sd = sd(value))
```

```{r}
# create a correlation plot
corrplot(cor(edu_data_scale), method = "number", type = "lower")
```

```{r}
# take only home school
diss_matrix <- edu_data_scale %>%
              select(home_school) %>%
              dist(method = "euclidean") 

```
```{r}
# what it's doing in the background, not needed 
matrix(1:12, 4)                           
dist(matrix(1:12, 4),method = "euclidean") # distance from each other
```


```{r}
# create a dissimilarity matrix showing distances
fviz_dist(diss_matrix)
```

## Building our dendrogram

```{r}
clusters <- diss_matrix %>%
                 hclust(method = "complete")
```

```{r}
clusters %>%
  plot(cex = 0.7, hang = .1)
```

```{r}
alt_clusters <- diss_matrix %>%
                    hclust(method = "ward.D2")
```


```{r}
clustering_dendrogram <- clusters %>%
                                  as.dendrogram() %>%
                                  dendextend::set("labels_cex", 0.5) 

clustering_dendrogram
```


```{r}
alt_clustering_dendrogram <- alt_clusters %>%
                              as.dendrogram() %>%
                                 dendextend::set("labels_cex", 0.6)

alt_clustering_dendrogram
```


```{r}
dend_diff(clustering_dendrogram, alt_clustering_dendrogram) 
```

The red lines show diffrences 

```{r}
plot(clusters, cex = 0.6, hang = -1)
```

```{r}
plot(clusters, cex = 0.6, hang = -1)
rect.hclust(clusters, k = 2, border = 2:5)
```

This looks not too bad. Groups don’t seem too big, or too small. In this case, I’d be tempted to stop here and say something along the following lines:

> We investigated education rates (per 100000) across US States using cluster analysis. Based on this analysis, we have defined two cluster groupings. States such as Alaska and Kentucky have similar education rates to one another. States such as Vermont and Maine have similar education rates to one another. These states have different education rates to one another. From our analysis, we could use these clusters to predict future education rate data in these states.


```{r}
# not needed

# This is base r palette, used above in border =
palette()
```

```{r}
edu_clustered_h2 <- edu_data %>%
                          mutate(school_cluster = cutree(clusters, k=2))

head(edu_clustered_h2,20)
```



Task - 20 mins

>  Spend some time carrying out cluster analysis for your other variables. What do you find for each variable? How many clusters do you choose, and why? What are your interpretations from your analysis?

>   What happens if you include all your data and perform clustering? (i.e. do not subset the data as we did in one of our first steps). For this, you have moved from predicting a specific variable to predicting something you might denoted as “overall education rates”. Do you get different results when the clustering takes into account all your data at once for each state? How would you interpret your results?

### Overall Education Rates

```{r}
overall_matrix <- edu_data_scale %>%
              dist(method = "euclidean") 
```

```{r}
overall_clusters <- overall_matrix %>%
                 hclust(method = "complete")
```

```{r}
overall_alt_clusters <- overall_matrix %>%
                    hclust(method = "ward.D2")
```


```{r}
overall_clustering_dendrogram <- overall_clusters %>%
                                  as.dendrogram() %>%
                                  dendextend::set("labels_cex", 0.5) 

overall_clustering_dendrogram
```

```{r}
overall_alt_clustering_dendrogram <- overall_alt_clusters %>%
                              as.dendrogram() %>%
                                 dendextend::set("labels_cex", 0.6)

overall_alt_clustering_dendrogram
```

```{r}
dend_diff(overall_clustering_dendrogram, overall_alt_clustering_dendrogram) 
```

```{r}
plot(overall_clusters, cex = 0.6, hang = -1)
rect.hclust(overall_clusters, k = 2, border = 2:5)
```

### State School

```{r}
state_matrix <- edu_data_scale %>%
              dist(method = "euclidean") 
```

```{r}
state_clusters <- state_matrix %>%
                 hclust(method = "complete")
```

```{r}
state_alt_clusters <- state_matrix %>%
                    hclust(method = "ward.D2")
```


```{r}
state_clustering_dendrogram <- state_clusters %>%
                                  as.dendrogram() %>%
                                  dendextend::set("labels_cex", 0.5) 

state_clustering_dendrogram
```

```{r}
state_alt_clustering_dendrogram <- state_alt_clusters %>%
                              as.dendrogram() %>%
                                 dendextend::set("labels_cex", 0.6)

state_alt_clustering_dendrogram
```

```{r}
dend_diff(state_clustering_dendrogram, state_alt_clustering_dendrogram) 
```

```{r}
plot(state_clusters, cex = 0.6, hang = -1)
rect.hclust(state_clusters, k = 2, border = 2:5)
```

### Private School

```{r}
private_matrix <- edu_data_scale %>%
              dist(method = "euclidean") 
```

```{r}
private_clusters <- private_matrix %>%
                 hclust(method = "complete")
```

```{r}
private_alt_clusters <- private_matrix %>%
                    hclust(method = "ward.D2")
```


```{r}
private_clustering_dendrogram <- private_clusters %>%
                                  as.dendrogram() %>%
                                  dendextend::set("labels_cex", 0.5) 

private_clustering_dendrogram
```

```{r}
private_alt_clustering_dendrogram <- private_alt_clusters %>%
                              as.dendrogram() %>%
                                 dendextend::set("labels_cex", 0.6)

private_alt_clustering_dendrogram
```

```{r}
dend_diff(private_clustering_dendrogram, private_alt_clustering_dendrogram) 
```

```{r}
plot(private_clusters, cex = 0.6, hang = -1)
rect.hclust(private_clusters, k = 2, border = 2:5)
```

After clusters have been created often you will want to give them labels to give them meaning and context (as discussed earlier today).

> Remember, dendrograms are useful tools, but are far from perfect. Because of they way in which we only join the two closest clusters at each step, we may find that a data point is occasionaly placed in the “wrong” cluster. This is only generally a problem when dealing with points which are very close together, however. In general a mis-sorted cluster will be grouped back in with its “correct” cluster at a later point in the algorithm, so assuming we are looking for a sufficiently small number of clusters the problem should resolve itself.


# john's graphs to better read the cluster
```{r}
edu_data %>% 
  rownames_to_column('state') %>% 
  mutate(school_cluster = cutree(clusters, 4)) %>% 
  ggplot(aes(home_school, state_school, colour = factor(school_cluster))) +
  geom_point() +
  geom_text(aes(label = state), check_overlap = T, size = 3,
            hjust = 'inward', vjust = 'inward') +
  guides(colour = 'none')
```

```{r}
edu_data %>% 
  rownames_to_column('state') %>% 
  mutate(school_cluster = cutree(clusters, 3)) %>% 
  ggplot(aes(private_school, state_school, colour = factor(school_cluster))) +
  geom_point() +
  geom_text(aes(label = state), check_overlap = T, size = 3,
            hjust = 'inward', vjust = 'inward') +
  guides(colour = 'none')
```
