---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


```{r}
actors <- read_csv("data/actors.csv")
movies <- read_csv("data/movies.csv")
roles <- read_csv("data/roles.csv")
```

```{r}
names(actors)
names(movies)
names(roles)

roles
```

Question 2.

    What are the primary keys and the foreign keys for each data frame?
    What kind of relationship (one-to-one, one-to-many, many-to-many) is there between
        actors and roles    # many-to-many
        movies and roles    # one-to-many
        actors and movies   # one-to-one

#2
Question 1.
From the movies table, find the id of the film The Little Mermaid. (In the data as “Little Mermaid, The”).
```{r}
movies %>% 
  filter(name == "Little Mermaid, The")
```

Question 2.
Filter the roles data frame, so we just have roles from the Little Mermaid (hint: use the ID you found in the last question). Save the result as little_mermaid_roles.
```{r}
little_mermaid_roles <- roles %>% 
  filter(movie_id == 192017)
little_mermaid_roles
```
Question 3.
Carry out a left join on this new data frame and the actors data frame, to find the actors which appeared in The Little Mermaid.

```{r}
left_join(little_mermaid_roles, actors, by = c("actor_id" = "id"))
```

Question 4.
Do the last join, but with an inner join instead. Is the data different? Why is that?
```{r}
inner_join(little_mermaid_roles, actors, by = c("actor_id" = "id"))
```

#3

Question 1.
For all actors who have appeared in at least two films. Call this data frame actors_more_than_once.
```{r}
actors_more_than_once <- actors %>% 
  filter(film_count >= 2)

actors_more_than_once
```

Question 2.
Left join actors_more_than_once to roles, so we can see all the actors roles those actors played in.
```{r}
left_join(actors_more_than_once, roles, by = c("id" = "actor_id"))
```



