---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidytext)
library(tidyverse)
```

```{r}
phrases <- c(
  "here is some text",
  "again more text",
  "text is text"
)
```

```{r}
example_text <- tibble(id = seq(phrases),
                       phrase = phrases)

example_text
```

```{r}
words_df <- example_text %>% 
  unnest_tokens(word, phrase)

words_df
```

```{r}
words_df %>% 
  count( word, id, sort =  TRUE,  name = "count")
```

```{r}
phrases <- c(
  "Here is some text.",
  "Again, more text!",
  "TEXT is text?"
)

example_text <- tibble(
  phrase = phrases,
  id     = 1:3
)

example_text %>%
  unnest_tokens(word, phrase) # auto lowercase and takes out punctuations

example_text %>%
  unnest_tokens(word, phrase, to_lower = FALSE, strip_punct = FALSE)
```
Task - 5 minutes

Below is the first four lines of the Robert Frost poem “Stopping by Woods on a Snowy Evening”.

```{r}
lines <- 
c(
  "Whose woods these are I think I know.",
  "His house is in the village though;", 
  "He will not see me stopping here",
  "To watch his woods fill up with snow."
)
```


    Create a data frame that has two variables: one with each word, the second with the line number of the word.
    Use this data frame to find all the words that appear more than once in the four lines.

```{r}
task_text <- tibble(
  phrase = lines,
  line    = 1:4
)

task_text %>%
  unnest_tokens(word, phrase) 
```


```{r}

# not in the notes 
crate_text_df <- function(text) {
  tibble(line_no = seq(text), line = text) %>% 
  unnest_tokens(output = word, input = line)
}

crate_text_df(lines) # can run lines or phrases
```

```{r}
library(janeaustenr)
```

```{r}
prideprejudice
```

```{r}
# using the function we made above that's not in the notes line 85

pride_book <- crate_text_df(prideprejudice)

pride_book
```

```{r}
stop_words
```

```{r}
stop_words %>% 
  group_by(lexicon) %>% 
  arrange(word) %>% 
  slice(1:5)
```

```{r}
pride_book %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE)
```

Task - 5 minutes

Find the most common words, not including stop words, in the book “Sense and Sensibility”.

```{r}

sense_text <- crate_text_df(sensesensibility)


sense_text %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  slice(1:5)
```


```{r}
# the way in the notes without the function we made

sense_book <- tibble(
  id = 1:length(sensesensibility),
  text = sensesensibility
) %>%
  unnest_tokens(word, text)

sense_book %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE)
```

