---
title: "Control Flow"
output: html_notebook
---

# Control Flow

## If Statemants


```{r}
if (condition) {
  # code that runs if the condition is met
}

# there can be a space between the if ()
# next thing should be placed on the next line
```

## Conitional Operator

`==` is equal to
`!=` not equal to
`<` less than
`>=` greater than OR equal to
`%in%` is in the set

```{r}
if (5 > 3) {
  print("yes")
}
```


```{r}
if (condition) {
  # do something
  
  #calculate something
  
  #
}

#will always be executed as it's outside the if
```


```{r}
x <- 7
y <- 8

if (x<= y) {
  print("it's less than or equal to y")
}
```
# Task
Write an if statement that prints 'Yes that divides by 2' if a number x is divisible by 2 (hint: use %% 2 == 0 as your expression), try it out for different values of x.

```{r}
x <- 4

if (x %% 2 == 0) {
  print("Yes the divides by 2")
}
```

## Nested Ifs

We can write more complicated if statments

```{r}
if (10 != 5){
  if (2 > 0) {
    print("the second condition is met")
  }
  print("the first condition is met")
}
```
```{r}
if (10 != 5){
  if (2 == 0) {
    print("the second condition is met")
  }
  print("the first condition is met")
}
```

Task 
1. `x` is 10 and `y` is 3
2. `x` is 10 and `y` is 2
3. `x` is 2 and `y` is 10

```{r}
x <- 2
y <- 10

if (x > y) {
  print("Start of outer if block")

  if (x %% y == 0) {
    print("Inner if block 1")
  }

  if (y %in% c(1:5)) {
    print("Inner if block 2")
  }

  print("End of outer if block")
}

print("After if statement")
```


## If... Else

```{r}
if (condition) {
  # do something
} else {
  # do something else
}
```

```{r}
if (rain) {
  waer_wellies()
} else {
  wear_trainers()
}
# if it's raining wear wellies, otherwise wear trainers
```

```{r}
if (4 < 5) {
  print("yes, it's less")
} else {
  print("no, it's not")
}
```

```{r}
if (2 %in% 50:100) {
  print("it's in the set")
} else {
  print("it's not in the set")
}
```
Task - 5 minutes
Edit your code from the first exercise so that it prints ‘Yes that divides’ when x is divisible by 2, and prints ‘No that does not divide’ otherwise.

```{r}
x <- 11

if (x %% 2 == 0) {
  print("yes that divides")
} else {
  print("no that doesn't divde")
}
```

