---
title: "Data Structures"
output: html_notebook
---

```{r}
nums <- 1:5
strings <-  c("I", "am", "text")
booleans <-  c(TRUE, FALSE)

class(nums)
```

# Lists

create them with `list()` #`` backwards quotes to make it code when kitied

```{r}
list("text", TRUE, c(1, 2, 3))
```

`as.list`
```{r}
as.list(nums)
```


```{r}
x <- list(1, "a", TRUE, c(1, 2, 3))

x[[2]]
x[2]

class(x[[2]]) # this truns it back into a vector 
class(x[2])   # this keeps it as a list
```

`[]` will always retrun the same data structure


## Naming Lists

```{r}
named_list <- list(a = "I love R", b=1:10)

named_list
```

```{r}
named_list$b
```

```{r}
named_list[["a"]]
```

```{r}
returned_value <- "a"

named_list[[returned_value]]
```

```{r}
named_list <- list(a = "I love R", b=1:10)

named_list[[2]]
```

## Nested List

You can put a list inside another list (inside another list)

```{r}
student_info <- list(
    "Agnes" = list(height = 1.8, shoe_size = 8L),
    "Bobby" = list(height = 2.1, gender = "Male"),
    "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info
```

```{r}
student_info <- list(
    "Agnes" = list(height = 1.8, shoe_size = 8L),
    "Bobby" = list(height = 2.1, gender = "Male"),
    "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info$Bobby$height
```

```{r}
student_info <- list(
    "Agnes" = list(height = 1.8, shoe_size = 8L),
    "Bobby" = list(height = 2.1, gender = "Male"),
    "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info[["Agnes"]][["shoe_size"]]
```


```{r}
student_info[[1]][[2]]
```


Task - 5 minutes

Make a named list that contains a numeric vector and a character vector.

Access the second element of your list using double square brackets [[]]. What is it's class?

Access the second element of your list using single square brackets []. What is it's class?

Access the second element of your list by name. What is it's class?

```{r}
d <- list(a = 1:10, b = "same")

d[[2]]
class(d[[2]])

returned <- "b"

d[returned]
class(d[returned])

d$b
```

# Johny work
```{r}
# Question 1

example_list <- list(a = c(1,4,7,9,11), b=  c("J","M","L","R","P"))

example_list

# Question 2

example_list[[2]]

class(example_list[[2]])

# Question 3

example_list[2]

class(example_list[2])

# Question 4

example_list[["b"]]

class(example_list[["b"]])
```


# Matrices
Will populate matrices by cols by default

```{r}
matrix(1:4, nrow = 2,ncol = 2)
```
To populate matrices by row, specify `byrow = TRUE`

```{r}
matrix(1:4, nrow = 2, ncol = 2, byrow = TRUE)
```

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
```

```{r}
class(m)
```

```{r}
as.data.frame(m)
```


# Selecting elements in a matrix

m[row, colum]

```{r}
# 1st row, 3rd colum
m[1,3]
# the first row
m[1,]
# 3rd colum
m[ , 3]
```

You can select multiple rows or colums by passing a vector

```{r}
m[, c(1, 3)]
# have picked 2 colums and dropped the middle one
```

```{r}
m * 5
# timesed all the matrix by 5
```
# This can be use in dplyr as well
```{r}
library(dplyr)
library(magrittr)

numers <- iris %>% select(-Species)

# if it's all the same data type
numers * 3

# access the first row of a df
numers[1, ]
```

## WOrking with rows and colums

- want to get the mean average of a column of a matrix

As we're out of the tidyveres we don't need to follow snake case

`colMeans()`

```{r}
colMeans(m)
```

Other convenience functions:

`rowSums`
`colSums`
`rowMeans`
`colMeans`

```{r}
rowSums(m)
```

These functions also work on dfs, as long as the columns are the same type

```{r}
colSums(numers)
```

```{r}
# you can specify the names of the columns, this over rides the vetor 
colnames(m) <- c("1st", "2nd", "3rd")

colSums(m)
```

Task - 5 minutes
Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
Find the element in the 2nd row, 3rd column.
Check that all the columns are identical by checking that their sums are the same.

```{r}
#Q1
p <- matrix(1:3, nrow = 3,ncol = 4)
p

#Q2
p[2, 3]

#Q3
colSums(p)
```



