---
title: "Tidy Data Mini-Lab - Answers"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

**Duration: 30 minutes** 

Now it's your turn to practice the verbs on a new dataset. Complete the following task:

# MVP

**Task 1.**  
Load in the `country_data.csv` file and take a look at it. What needs changed in this to make it tidy? Which columns need made into long format, and which need spread to rows?

<details>
<summary>**Answer**</summary>
```{r, warning = FALSE, message= FALSE}
library(tidyverse)

# read in the data 
country_data <- read_csv("data/country_data.csv")
```
</details>
<br>

**Task 2.**  
Do the appropriate pivoting of your columns using R  

<details>
<summary>**Answer**</summary>
```{r}
# gather the year columns into a new one called year
country_long <- country_data %>%
  pivot_longer(cols =  starts_with("Y"), 
               names_to = "year", 
               values_to = "visitors")

country_long
```

```{r}
# spread the case type into two columns
country_wide <- country_long %>%
  pivot_wider(names_from = type, 
             values_from = visitors)

country_wide
```
</details>
<br>

**Task 3.**  
Separate the `country` variable so you have the name in one column and the number in another.   

<details>
<summary>**Answer**</summary>
```{r}
# separate the country variable so you only have the name 
country_separated <- country_wide %>%
  separate(country, c("country_n", "country_name"), sep = "\\. " )

country_separated
```
</details>
<br>

**Task 4.**  
Unite the `code_letter` and `code_numb` variables so that they are both in one column, with no spaces separating them.   

<details>
<summary>**Answer**</summary>
```{r, message = FALSE, warning = F}

# unite the code_letter and code_numb variables into a new one called country_code
country_united <- country_separated %>%
  unite(country_code, 
        c(code_letter, code_numb), 
        sep = "")

country_united
```
</details>

