---
title: "Tidy Data Mini-Lab"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

**Duration: 30 minutes** 

Now it's your turn to practice the verbs on a new dataset. Complete the following task:

# MVP

**Task 1.**  
Load in the `country_data.csv` file and take a look at it. What needs changed in this to make it tidy? Which columns need made into long format, and which need spread to rows?

```{r}
country_data <- read_csv("data/country_data.csv")
head(country_data)
```

<br>

**Task 2.**  
Do the appropriate pivoting of your columns using R  
```{r}
country_data_long <- country_data %>% 
  pivot_longer(
    cols = starts_with("Y"),
    names_to = "year",
    values_to = "count",
    names_prefix = "Y"
  )

country_data_long
```
```{r}
country_data_wider <- country_data_long %>% 
  pivot_wider(names_from = "type",
              values_from = "count")

country_data_wider
```

<br>

**Task 3.**  
Separate the `country` variable so you have the name in one column and the number in another.   

```{r}
country_data_nocoded <- country_data_wider %>% 
  separate(country,                         # the colum you want separated
           c("code_numb", "country"),       # the names of the new colum, can over write code_numb as                                                 there both the same 
           sep = "\\.")                     # where to separate, \ ignore the general rule of next thing

country_data_nocoded
```


<br>

**Task 4.**  
Unite the `code_letter` and `code_numb` variables so that they are both in one column, with no spaces separating them.   

```{r}
country_data_untite <- country_data_nocoded %>% 
  unite(code,                             # the new colum name
        c("code_numb", "code_letter"),    # the colums being united together
        sep = "")                         # for no space just use ""

country_data_untite
```



