---
title: "Spatail Data"
output: html_notebook
---


#  Spatail Data

What do you think spatial data looks like?
What do you think a spatail data vis looks like?
Can you find any exaample online?


- map-able data

spatial (geospatial) data - data about 



How are spatial data represented?

spatial data are specail - while encoding as numbers (double) - these numbers actually have special meaning

spatial vectors (non R-specific)

3 main types

- Point (1, 4) train station locations, wells

- Line (Line string) ((1, 4), (4, 5)) roads 

- Polygon  ((1, 4), (4, 5), ....) council areas, lakes


# Spatial Data in R

sometimes we just get given dataframes with lat-long colums

another option is to load in shapefiles

```{r}
library(sf) #  sf = simple fratures
library(tidyverse)

```

a feature = a gometry
```{r}
north_carolina <- st_read(system.file("shape/nc.shp", 
                                      package = "sf"))
```

```{r}
north_carolina %>% 
  as.tibble()
```

```{r}
nc_geo <- st_geometry(north_carolina)
```

```{r}
nc_geo[[1]]
```

```{r}
class(north_carolina)
```
it's a dataframe with geometries

# Plotting Geometries

```{r}
plot(north_carolina["AREA"])
```

`st_as_sf()` - convert non-sf objects (that include geometries) into sf objects

```{r}
# subset the county 
plot(north_carolina[1,1])
```

### Task - 5 mins

Have a look through some of the variables within your north_carolina dataset, and see if you can create a spatial plot using the techniques above.

```{r}
plot(north_carolina[3, 3, 1, 1])
```

```{r}
plot(north_carolina["PERIMETER"])
```

```{r}
plot(north_carolina[c(1, 2, 12), "AREA"])
```

# ggplot and sf

`geom_sf()`

```{r}
 north_carolina %>% 
  ggplot() +
  geom_sf(aes(fill = SID74), colour = "black") +
  theme_bw()
```

### Task : 5 mins

Try your hand at plotting another feature from the north_carolina dataset. What does it tell you?
```{r}
 north_carolina %>% 
  ggplot() +
  geom_sf(aes(fill = NWBIR74), colour = "yellow") +
  theme_gray()
```

```{r}
north_carolina %>% 
  ggplot() +
  geom_sf(aes(fill = BIR74), colour = "red", size = 0.1) +
  scale_fill_continuous(type = "viridis") +
  theme_classic()
```

```{r}
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
```

Using rnaturalearth package we can inport bou

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
```

```{r}
view(world)
head(world, 1)
```

```{r}
world %>% 
  as.tibble() %>% 
  head(5)
```

okay so we've got the data, now we can plot it

```{r}
world %>% 
  ggplot() +
  geom_sf() +
  labs(x = "longitude", y = "latitude", title = "World map")
```

We can plot actual data now

```{r}
world %>% 
  ggplot() +
  geom_sf(aes(fill = pop_est)) +
  scale_fill_viridis_c()
```

### Task - 5 mins

Recap your knowledge from ggplot week, and set your geom_sf aesthetic to be filled with the estimated gdp (gdp_md_est variable). Extra points if you make your map colour blind friendly!

What does your plot tell you? What does it tell you compared to the population?

```{r}
world %>% 
  ggplot() +
  geom_sf(aes(fill = gdp_md_est)) +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```

```{r}
world %>% 
  ggplot() +
  geom_sf(aes(fill = continent)) 
```

```{r}
world %>% 
  filter(continent != "Antarctica") %>% 
  ggplot() +
  geom_sf(aes(fill = continent)) 
```

# 4.3 Subsetting simple features

What if we only care about one particular country? We can subset the data as we would in tidyverse:

```{r}
country_italy <- world %>% 
  filter(name == "Italy")
```

```{r}
country_italy %>% 
  ggplot() +
  geom_sf() +
  labs(x = "Longitude", y = "Latitude", title = "Italia")
```

```{r}
country_scot <- world %>% 
  filter(name == "Scotland") 
  
```

```{r}
country_scot %>% 
  ggplot() +
  geom_sf() +
  labs(x = "Longitude", y = "Latitude", title = "Scotland")
```

```{r}
country_uk <- world %>% 
  filter(name == "United Kingdom") 
```

```{r}
country_uk %>% 
  ggplot() +
  geom_sf() +
  labs(x = "Longitude", y = "Latitude", title = "UK")
```

# Zooming in on particular parts of the world

- we can subset our graph by limiting the x and y range using `coord_sf()`
- ggplot

```{r}
world %>% 
  ggplot() +
  geom_sf() +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```


```{r}
ggplot(data = world) +
    geom_sf(aes(fill = pop_est)) +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```

Add informative lables

- we need to tell ggplot where to put the lables
- we need to calculate where we want to put the labels
(centre of each feature)
- we need to calculate the centres (centroids)

```{r}
world %>% 
  mutate(centre = st_centroid(st_make_valid(geometry))) %>% 
  ds_tibble() %>% 
  select(name, centre)
```

```{r}
world_centre <- world %>% 
  mutate(centre = st_centroid(st_make_valid(geometry))) %>% 
  mutate(lat = st_coordinates(centre)[,1],
         long = st_coordinates(centre)[,2])
```

```{r}
# plot these, along side our zoomed in 
ggplot(data = world_centre) +
geom_sf(fill ="grey") +
 geom_text(aes(x = lat , 
               y = long, 
               label = name), 
           color = "darkblue", 
           fontface = "bold", 
           check_overlap = TRUE) +
coord_sf(xlim = c(-102.15, -74.12), 
         ylim = c(7.65, 33.97), 
         expand = FALSE)
```

we can add additionla information using annotate

- add one that says: Gulf of Mexico

```{r}
# the annotate function can be used to add a single character string at a specific location, as demonstrated here to add the Gulf of Mexico. 
ggplot(data = world_centre) +
geom_sf(fill ="grey") +
 geom_text(aes(x = lat , y = long, 
               label = name), 
           color = "darkblue", 
           fontface = "bold", 
           check_overlap = TRUE) +
annotate(geom = "text", x = -90, y = 26, 
         label = "Gulf of Mexico", 
         fontface = "italic", 
         color = "grey22", 
         size = 6) +
coord_sf(xlim = c(-102.15, -74.12), 
         ylim = c(7.65, 33.97), 
         expand = FALSE)
```

### Task - 10 mins

Have a play around with the plot above, and see what changing parameters does, what using different variables does, what else you can plot, etc

```{r}
ggplot(data = world_centre) +
geom_sf(fill ="grey") +
 geom_text(aes(x = lat , y = long, 
               label = name), 
           color = "darkblue", 
           fontface = "bold", 
           check_overlap = TRUE) +
  annotate(geom = "point", x = -88, y = 28, 
           label = "Mad Dog field", 
           fontface = "italic", 
           color = "black", 
           size = 4) +
  annotate(geom = "text", x = -88, y = 27, 
           label = "Mad Dog field", 
           fontface = "italic", 
           color = "black", 
           size = 4) +
  annotate(geom = "point", x = -88, y = 30, 
           label = "moblile", 
           fontface = "italic", 
           color = "black", 
           size = 4) +
  annotate(geom = "text", x = -88, y = 29.5, 
           label = "Moblile", 
           fontface = "italic", 
           color = "black", 
           size = 4) +
  annotate(geom = "text", x = -90, y = 24, 
           label = "Gulf of Mexico", 
           fontface = "italic", 
           color = "grey22", 
           size = 6) +
  coord_sf(xlim = c(-102.15, -74.12), 
           ylim = c(7.65, 33.97), 
           expand = FALSE)
```



projectioon; wgs84 (1984)

####Louise work
```{r}
world_centre %>%
  ggplot() +
  geom_sf(aes(fill = income_grp)) +
  geom_text(aes(x = lat, y = long, label = name), colour = "black",
            fontface = "bold", check_overlap = TRUE, size = 3) +
  annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", size = 3,
           fontface = "italic")+
  coord_sf(xlim = c(-110.15, -60.12), ylim = c(2.65, 50.97), expand = TRUE) +
  labs(x = "Latitude",
       y = "Longitude",
       fill = "OECD Income group")
```

#### Ricardo work
```{r}
world_centre %>% 
  ggplot() +
  geom_sf(aes(fill = name), show.legend = F) +
  geom_text(aes(lat, long, label = name), size = 2.5) +
  annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico",
           size = 4, fontface = "italic",
           check_overlap = TRUE) +
  coord_sf(xlim = c(-110, -60), ylim = c(8, 35), expand = T) +
  theme_minimal() 
```


```{r}
world_centre %>% 
  ggplot() +
  geom_sf(aes(fill = name), show.legend = F) +
  geom_text(aes(lat, long, label = name), size = 2.5) +
  annotate(geom = "text", x = 5, y = 57.5, label = "North Sea",
           size = 4, fontface = "italic",
           check_overlap = TRUE) +
  coord_sf(xlim = c(02, 010), ylim = c(56, 63), expand = T) +
  theme_minimal()
```


```{r}
library(leaflet)
```

```{r}
leaflet() %>% 
  addTiles() %>% #basemap
  addMarkers(lng = 174.768, lat = -36.852, popup = "The birth place of R")
```

That was our first leaflet

get some spatial data from the web
- turn it into a format R can work with
- visualise it using leaflet

```{r}
library(jsonlite)
```
```{r}
colarado_data_url <- "https://data.colorado.gov/resource/j5pc-4t32.json?&county=BOULDER"
```


```{r}
head(readLines(colarado_data_url))
```

jsonlite package has a lot of functions to help work with json data

json is like a list of lists in R - sometimes we need to recursively sift through to extract the relevant data

```{r}
water_data_df <- fromJSON(colarado_data_url) %>%
  # remove the nested data frame 'location' column to unnest the latitude and longitudes using 'flatten' from jsonlite
  jsonlite::flatten(recursive = TRUE) 
```

```{r}
water_data_df_clean %>% 
  select(-location.needs_recoding) %>% 
  mutate(across(starts_with("location"), as.numeric) %>% 
  filter(!is.na(location.latitude), !is.na(location.latitude)))
```



```{r}
# turn columns to numeric and remove NA values
water_data_df_clean <- water_data_df %>%
  mutate_at(vars(amount, location.latitude, location.longitude), as.numeric) %>%
  filter(!is.na(location.latitude))

# create leaflet map
leaflet(water_data_df_clean) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~location.longitude,
                   lat = ~location.latitude)
```

- incidences of surface water in abd around Boulder Coloardo


```{r}
leaflet(water_data_df_clean) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~location.longitude,
                   lat = ~location.latitude,
                   radius = ~ sqrt(amount)/100)
```


```{r}
leaflet(water_data_df_clean) %>%
  addTiles() %>%  
  addMarkers(lng = ~location.longitude,
                   lat = ~location.latitude)
```

we can add clustering options

```{r}
leaflet(water_data_df_clean) %>%
  addTiles() %>%  
  addMarkers(lng = ~location.longitude,
                   lat = ~location.latitude,
             clusterOptions = markerClusterOptions())
```

# Leaflet in Shiny

```{r}
library(leaflet)
library(CodeClanData)
```

```{r}
leaflet(whisky) %>% 
  addTiles() %>%
  addCircleMarkers(lat = ~Longitude, lng = ~Latitude, popup = ~Distillery)
```







