---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1.** 

Read in the data and investigate it, e.g. variables, variable types, dimensions etc. df -> minilab

```{r}
library(tidyverse)
```

```{r}
minilab <- read_csv("4_extra_dplyr_practice/data/wine_ratings.csv")
```

```{r}
class(minilab)
```

```{r}
dim(minilab)
```

```{r}
summary(minilab)
```
```{r}
names(minilab)
```

```{r}

```





**Question 2.** 

Select columns `title`, `taster_name`, `points`, `price`, and `country`, then **drop all NAs**. Save into a variable called `wine_subset`. Use this variable for all subsequent tasks.
```{r}
wine_subset <- minilab %>% 
  select(title, taster_name, points, price, country)
  
wine_subset
```

```{r}
wine_subset %>% 
summarise(across(everything(), ~ sum(is.na(.))))
```

```{r}
wine_subset %>% 
  drop_na() %>% 
  summarise(across(everything(), ~ sum(is.na(.))))


```

**Question 3.** 

Find the 5 countries with the highest mean points.

```{r}
wine_subset %>%
  #dplyr::select(title, taster_name, points, price, country) %>% 
  group_by(title) %>% 
  mutate(mean_points = mean(points)) %>% 
  arrange(desc(mean_points)) %>% 
  head(5)
```

**Question 4.** 

Find the 5 countries with the lowest mean prices.

```{r}
wine_subset %>%
  dplyr::select(title, taster_name, points, price, country) %>% 
  group_by(title) %>% 
  mutate(mean_points = mean(points)) %>% 
  arrange(desc(mean_points)) %>% 
  tail(5)
```

**Question 5.** 

Does the result differ if you use the median instead of the mean for Questions 3 and 4?

```{r}
wine_subset %>%
  dplyr::select(title, taster_name, points, price, country) %>% 
  group_by(title) %>% 
  mutate(median_points = median(points)) %>% 
  arrange(desc(median_points)) %>% 
  head(5)
```

```{r}
wine_subset %>%
  dplyr::select(title, taster_name, points, price, country) %>% 
  group_by(title) %>% 
  mutate(median_points = median(points)) %>% 
  arrange(desc(median_points)) %>% 
  tail(5)
```

**Question 6.** 

Find the names of all the tasters using `distinct`. Then take a random sample of 5 names using `sample_n`. Save these 5 names into a new variable called `taster_sample` which should be a **character vector** containing only the names. 

```{r}
taster_sample <- wine_subset %>% 
  distinct(taster_name) %>% 
  sample_n(5)

taster_sample
```

**Question 7**

Do the same thing for `country`, i.e. create a **character vector** called `country_sample` using `distinct` and `pull`, but this time use `sample_frac` to pick 10 countries at random. 

```{r}
country_sample <- wine_subset %>% 
  distinct(country) %>% 
  sample_n(10)

country_sample
```

**Question 8**

Filter the rows of `wine_subset` so you are left only with data from the 5 tasters in your `taster_sample` and from the countries in your `country_sample`. Save this into a new variable called `wine_sample`.

```{r}
wine_sample <- filter(wine_subset, )
```


