---
title: "Facets"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)

chick_weight <- clean_names(ChickWeight)

ggplot(chick_weight) + 
  aes(x = time, y = weight, group = chick, colour = diet) +
  geom_line() +
  geom_smooth(aes(group = diet), colour = "black")
```


```{r}
ggplot(chick_weight) + 
  aes(x = time, y = weight, group = chick, colour = diet) +
  geom_line() +
  geom_smooth(aes(group = diet), colour = "black") +
  facet_wrap(~ diet)
```


```{r}
ggplot(chick_weight) + 
  aes(x = time, y = weight, group = chick, colour = diet) +
  geom_line() +
  geom_smooth(aes(group = diet), colour = "black") +
  facet_wrap(~ diet, nrow = 1)
```

```{r}
benefits <- read_csv("clean_data/benefits.csv")
```


### Task Use geom_histogram to create a histogram showing the number of unemployed by age. Then facet the graph based on biological sex.
```{r}
ggplot(benefits) +
  geom_histogram(aes(x = age, fill = sex)) +
  facet_wrap(~sex, ncol = 2)
```


```{r}
ggplot(benefits) +
  geom_histogram(aes(x = age, fill = sex), col = "white") +
  facet_wrap(~sex, ncol = 2)
```

```{r}
ggplot(benefits) + 
geom_histogram(aes(x = age, fill = sex), col = "white", alpha = 0.5, position = 'identity') #+ 
#facet_wrap( ~ sex, ncol = 1)
```


```{r}
ggplot(benefits) + 
geom_histogram(aes(x = age, fill = sex), col = "white", alpha = 0.5)
```


```{r}
ggplot(benefits) + 
  geom_histogram(aes(x = age, fill = sex), col = "white") + 
  facet_wrap(married ~ sex)

ggplot(benefits) + 
  geom_histogram(aes(x = age, fill = sex), col = "white") + 
  facet_grid(married ~ sex)
```


```{r}
benefits %>% 
  mutate(married_sex = case_when(married == 'yes' ~ paste0('Married\n ', sex),
                                 married == 'no' ~ paste0('Unmarried ', sex))) %>% 
ggplot() + 
  geom_histogram(aes(x = age, fill = sex), col = "white",
                 show.legend = FALSE) + 
  # geom_text(aes(60, 150, label = married_sex), hjust = 1) +
  facet_wrap(~ married_sex)
```

```{r}
benefits %>% 
  mutate(married_sex = case_when(married == 'yes' ~ paste0('Married ', sex),
                                 married == 'no' ~ paste0('Unmarried ', sex))) %>% 
ggplot() + 
  geom_histogram(aes(x = age, fill = sex), col = "white",
                 show.legend = FALSE) + 
   geom_text(aes(60, 150, label = married_sex), hjust = 1) +
  facet_wrap(~ married_sex) +
  theme(strip.text.x = element_blank())
```

### Task Use the same dataset, but base the facet_grid() on marital status (married) and if the individual has children (dkids).
```{r}
benefits %>% 
  ggplot() +
  geom_histogram(aes(x = age, fill = sex), col = "white") +
  facet_grid(married ~ dkids)
  
```

### Extension: can you improve the way that the facets are labled.
```{r}
benefits %>% 
  ggplot() +
  geom_histogram(aes(x = age, fill = sex), col = "white") +
  facet_grid(married ~ dkids, labeller = "label_both")
```

Louise's work
```{r}
ggplot(benefits) +
  geom_bar(aes(x = married, fill = dkids), col = "white") +
  facet_grid(married ~ dkids)
```

```{r}
students <-  read_csv("clean_data/students.csv")
```

```{r}
students %>% 
  ggplot(aes(reaction_time)) +
  geom_histogram()
```

gives us a warning `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.

```{r}
students %>% 
  ggplot(aes(reaction_time)) +
  geom_histogram(bins = 10)
```

