---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(tidyverse)
```


```{r}
ggplot(students) +
  geom_bar(aes(x = superpower))
```

```{r}
olympics_overall_medals <- read_csv("clean_data/olympics_overall_medals.csv")
```

```{r}
top_10 <- olympics_overall_medals %>%
  filter(season == "Summer", medal == "Gold") %>%
  arrange(desc(count)) %>%
  top_n(10)

top_10
```
### Task Create an informative plot that plots the count of medals by team. Write down an explanation of what the plot shows.


```{r}
ggplot(top_10) +
  geom_col(aes(x = team, y = count, fill = team)) + 
  xlab("Countries") +
  ylab("Num. of Medals") +
  ggtitle("Top 10 Counties with the most Golds")

ggplot(top_10) +
  geom_col(aes(x = count, y = team, fill = team)) + 
  xlab("Countries") +
  ylab("Num. of Medals") +
  ggtitle("Top 10 Counties with the most Golds")
```

```{r}
ggplot(olympics_overall_medals) +
  geom_bar(aes(x = team, fill = team, stat='identity')) + 
  xlab("Countries") +
  ylab("Num. of Medals") +
  ggtitle("Top 10 Counties with the most Golds")
```


```{r}
library(janitor)
```

```{r}
chick_weight <- clean_names(ChickWeight)

head(chick_weight)
```

```{r}
ggplot(chick_weight) +
  geom_line(
    aes(x = time, y = weight, group = chick, colour = diet))
```


```{r}
# lines first then points
ggplot(chick_weight) +
  geom_line(aes(x = time, y = weight, group = chick, colour = diet)) +
  geom_point(aes(x = time, y = weight, colour = diet))

# points first then lines
ggplot(chick_weight) +
  geom_point(aes(x = time, y = weight, colour = diet)) +
  geom_line(aes(x = time, y = weight, group = chick, colour = diet)) 
  

```

```{r}
ggplot(chick_weight) +
  geom_line(
    aes(x = time, y = weight, group = chick, colour = diet),
    alpha = 0.25
  ) +
  geom_point(
    aes(x = time, y = weight, colour = diet),
    alpha = 0.5
  ) +
  geom_smooth(
    aes(x = time, y = weight, group = diet), colour = "blue")

```

can write it as below if only using the global data, makes it netter 
```{r}
ggplot(chick_weight,
       aes(x = time, y = weight,
           group = chick, colour = diet)) +
  geom_line() +
  geom_point() +
  geom_smooth(aes(group = diet), colour = "black") +
  labs(colour = "DIET")
```


```{r}
ggplot(chick_weight) +
  geom_line(
    aes(x = time, y = weight, group = chick, colour = diet),
    alpha = 0.25
  ) +
  geom_point(
    aes(x = time, y = weight, colour = diet),
    alpha = 0.5
  ) +
  geom_smooth(
    aes(x = time, y = weight), colour = "blue")
  
```

```{r}
students <-  read_csv("clean_data/students.csv")
```

### Task
### Use geom_point to make a scatter graph, with the height of students on the x-axis and their reaction time of the y axis.
```{r}
ggplot(students) +
  geom_point(
    aes(x = height_cm, y = reaction_time)
  )
```

### Make all the points blue. For geom_bar, the colour of the bar is controlled by fill, but for geom_point the colour of the points are controlled by colour.
```{r}
ggplot(students) +
  geom_point(
    aes(x = height_cm, y = reaction_time), colour = "blue")
  
```


### Make the colour of the points depend on the superpower the student wishes they had.
```{r}
ggplot(students) +
  geom_point(aes(x = height_cm, y = reaction_time, colour = superpower))
```


### Write down what the graph tells you overall.
There doesn’t seem to be much of a relationship between height and reaction type.


```{r}
pets <- read_csv("clean_data/pets.csv")
```
### Task
### We want age of the x-axis and weight on the y axis
```{r}
ggplot(pets) + 
  aes(x = age, y = weight)
```

### We want the points the be different colours depending on the gender of the pet
```{r}
ggplot(pets) + 
  aes(x = age, y = weight) +
  geom_point(aes(colour = sex))
```

### We want different shapes depending on the type of animal
```{r}
ggplot(pets) + 
  aes(x = age, y = weight) +
  geom_point(aes(colour = sex, shape = animal))
```

### We want all the points to be bigger than normal (size 4).
```{r}
ggplot(pets) + 
  aes(x = age, y = weight) +
  geom_point(aes(colour = sex, shape = animal), size = 4)
```

### We also want labels with the pets names next to every point.
```{r}
ggplot(pets) + 
  aes(x = age, y = weight) +
  geom_point(aes(colour = sex, shape = animal), size = 4) +
  geom_text(
    aes(label = name),
    nudge_x = 0.5,
    nudge_y = 0.1,
  )
```

can use `hjust` to have the names move in towards the middle 
```{r}


ggplot(pets) + 
  aes(x = age, y = weight) +
  geom_point(aes(colour = sex, shape = animal), size = 4) +
  geom_text(
    aes(label = name),
    hjust ="inward"
  )

```

