---
title: "Two-sample hypothesis tests"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

```{r}
nice <- read_csv("data/nice.csv")%>% 
  clean_names()
algarve <- read_csv("data/algarve.csv") %>% 
  clean_names()
```

```{r}
apart_prices <- bind_rows(nice,algarve, 
                          .id = "location") %>% 
  mutate(location = if_else(location == "1", 
                            "nice", 
                            "algarve")) %>% 
  select(-x1)


head(apart_prices)
```

```{r}
apart_prices %>% 
  group_by(location) %>% 
  summarise(n = n())
```

```{r}
apart_prices %>% 
  ggplot(aes(y = price, x = location)) +
  geom_boxplot()
```

```{r}
null_distribution <- apart_prices %>% 
  specify(price ~ location) %>% #it is the relationship between price and location we are testing. This can also be written like specify(response = price, explanatory = location) 
  hypothesize(null = "independence") %>% #the null hypothesis is there is no relationship i.e. they are independent
  generate(reps = 1000, 
           type = "permute") %>% 
  calculate(stat = "diff in means", 
            order = c("algarve", 
                      "nice")) #our sample stat is mean of algarve minus mean of nice, so this is the order we specify in the calculate step

head(null_distribution)
```

Step by step, this does the following:

    Selects the data we want to use

    The syntax price ~ location in the specify() call says that we are testing whether price varies with location, so just read the tilde ~ as ‘varies with’. We could also write this as specify(response = price, explanatory = location)

    When using the infer functions for the hypothesize() step, it is the hypothesize(null = 'independence) line that selects a permutation test. If we didn’t specify that generate(type = "permute") it would use type=permute by default as this is how it carries out a test for independence between two variables.

    It generates a distribution of 1000 samples

    It calculates the difference between the mean of the algarve and nice samples.


```{r}
observed_stat <- apart_prices %>% 
  specify(price ~ location) %>% 
  calculate(stat = "diff in means", 
            order = c("algarve", 
                      "nice"))
  
observed_stat
```

```{r}
null_distribution %>%
  visualise() +
  shade_pvalue(obs_stat = observed_stat, 
               direction = "right")
```

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, 
              direction = "right")

p_value
```



