---
title: "Two-sample hypothesis tests"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

```{r}
nice <- read_csv("data/nice.csv")%>% 
  clean_names()
algarve <- read_csv("data/algarve.csv") %>% 
  clean_names()
```

```{r}
apart_prices <- bind_rows(nice,algarve, 
                          .id = "location") %>% 
  mutate(location = if_else(location == "1", 
                            "nice", 
                            "algarve")) %>% 
  select(-x1)


head(apart_prices)
```

```{r}
apart_prices %>% 
  group_by(location) %>% 
  summarise(n = n())
```

```{r}
apart_prices %>% 
  ggplot(aes(y = price, x = location)) +
  geom_boxplot()
```

```{r}
null_distribution <- apart_prices %>% 
  specify(price ~ location) %>% #it is the relationship between price and location we are testing. This can also be written like specify(response = price, explanatory = location) 
  hypothesize(null = "independence") %>% #the null hypothesis is there is no relationship i.e. they are independent
  generate(reps = 1000, 
           type = "permute") %>% 
  calculate(stat = "diff in means", 
            order = c("algarve", 
                      "nice")) #our sample stat is mean of algarve minus mean of nice, so this is the order we specify in the calculate step

head(null_distribution)
```

Step by step, this does the following:

    Selects the data we want to use

    The syntax price ~ location in the specify() call says that we are testing whether price varies with location, so just read the tilde ~ as ‘varies with’. We could also write this as specify(response = price, explanatory = location)

    When using the infer functions for the hypothesize() step, it is the hypothesize(null = 'independence) line that selects a permutation test. If we didn’t specify that generate(type = "permute") it would use type=permute by default as this is how it carries out a test for independence between two variables.

    It generates a distribution of 1000 samples

    It calculates the difference between the mean of the algarve and nice samples.


```{r}
observed_stat <- apart_prices %>% 
  specify(price ~ location) %>% 
  calculate(stat = "diff in means", 
            order = c("algarve", 
                      "nice"))
  
observed_stat
```

```{r}
null_distribution %>%
  visualise() +
  shade_pvalue(obs_stat = observed_stat, 
               direction = "right")
```

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, 
              direction = "right")

p_value
```

### Task - 10 mins

We have two more datasets for similar apartment lets in Corfu and Florence. Frame and perform an independent two-sample test to answer the following question:

‘On average, is the price of 2-bedroom, 14-day holiday apartment lets in Florence significantly lower than that of comparable lets in Corfu?’


    Choose α

and frame your hypotheses before you see the data
You will need to format your data in a combined, tidy dataset prior to performing the test

```{r}
corfu <- read_csv("data/corfu.csv") %>% 
  clean_names()

florence <- read_csv("data/florence.csv") %>% 
  clean_names()
```

H0: μprice(Corfu)−μprice(Florence)=0
Ha: μprice(Corfu)−μprice(Florence)<0

conventional significance level α=0.05 and write our hypotheses

N.B. I've done this the wrong way around - should have done 

Ha: μprice(Florence)−μprice(Corfu)<0

Ending up doing 

Ha: μprice(Corfu)−μprice(Florence)>0


```{r}
flo_cor <- bind_rows(corfu, 
                     florence, 
                     .id = "location") %>% 
            mutate(location = if_else(location == "1", 
                                      "corfu", 
                                      "florence")) %>% 
  select(-x1)

head(flo_cor)
tail(flo_cor)
```

```{r}
flo_cor %>% 
  group_by(location) %>% 
  summarise(n = n())
```

```{r}
flo_cor %>% 
  ggplot(aes(y = price, x = location)) +
  geom_boxplot()
```
Florence significantly higher than that of comparable lets in Corfu from the boxgraphm, so will fail to reject H0


```{r}
flo_cor_null <- flo_cor %>%
  specify(price ~ location) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 1000, 
           type = "permute") %>% 
  calculate(stat = "diff in means", 
            order = c("corfu", 
                      "florence"))

head(flo_cor_null)
```

```{r}
stat_flo_cor <- flo_cor %>% 
  specify(price ~ location) %>%
  calculate(stat = "diff in means", 
            order = c("corfu", "florence"))
  
stat_flo_cor
```

```{r}
flo_cor_null %>% 
   visualise() +
  shade_p_value(obs_stat = stat_flo_cor, 
                direction = "left")
```

```{r}
p_value <- flo_cor_null %>%
  get_p_value(obs_stat = stat_flo_cor, 
              direction = "left")

p_value
```

The p-value very close to 1, which is above α=0.05, so we fail to reject H0: there is no significant evidence that the prices of lets in Florence are lower than those in Corfu. Let’s boxplot the prices by location for further insight.


# 4 Paired (aka dependent) samples

```{r}
books <- read_csv("data/ucla_textbooks_f18.csv")
```

```{r}
glimpse(books)
```

```{r}
books_diff <- books %>% 
  mutate(diff_new = bookstore_new - amazon_new) %>% 
  filter(!is.na(diff_new))

books_diff %>% 
  ggplot(aes(x = diff_new)) +
  geom_histogram(col = "white")
```

t looks like for many of the books the price difference is centred around zero but there are some significant outliers towards the right.

Let’s set up our hypotheses:

H0: μdiff_new=0
Ha: μdiff_new≠0 (!= 0)

```{r}
books_null <-  books_diff %>% 
  specify(response = diff_new) %>% 
  hypothesise(null = "point", mu = 0) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

```{r}
head(books_null)
```

```{r}
books_stat <- books_diff %>% 
  specify(response = diff_new) %>% 
  calculate(stat = "mean")

books_stat
```

```{r}
books_null %>%
  visualise() +
  shade_p_value(obs_stat = books_stat, 
                direction = "both")
```

```{r}
book_p <- books_null %>% 
  get_p_value(obs_stat = books_stat, 
              direction = "both")

book_p
```

### Task - 10 mins

Frame and perform a paired hypothesis test to answer the following question:

‘On average, are the prices of used course texts lower on Amazon than at the campus bookstore?’


Hints:

    It will help to start with a clear definition: diff_used = bookstore_used - amazon_used or diff_used = amazon_used - bookstore_used
    This will be a one-tailed test, so frame your hypotheses appropriately.
    
    
```{r}
diff_books <- books %>% 
  mutate(diff_used = bookstore_used - amazon_used) %>% 
  filter(!is.na(diff_used))

diff_books %>% 
  ggplot(aes(x = diff_used)) +
  geom_histogram(col = "white")
```

#######


# 5 Testing differences in proportions

```{r}
poll <-  read_csv("data/election_data.csv")
```
```{r}
head(poll)
```

```{r}
ggplot(poll, aes(x = vote, 
                 fill = gender)) +
  geom_bar(position = "fill") +
  coord_flip()
```

```{r}
poll_stat <- poll %>% 
  specify(gender ~ vote, 
          success = "woman") %>% 
  calculate(stat = "diff in props", 
            order = c("obama", 
                      "romney"))

poll_stat
```

```{r}
poll_null <- poll %>% 
  specify(gender ~ vote, 
          success = "woman") %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 1000, 
           type = "permute") %>% 
  calculate(stat = "diff in props", 
            order = c("obama", 
                      "romney"))

head(poll_null)
```

```{r}
poll_null %>% 
  visualise() +
  shade_p_value(obs_stat = poll_stat,
                direction = "both")
```

```{r}
poll_p_value <- poll_null %>%
  get_p_value(obs_stat = poll_stat, 
              direction = "both")

poll_p_value
```


```{r}
ggplot() +
  geom
```

