---
title: "Some More Functions"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Examples of functions

Here's some examples of functions we've seen before and their inputs and outputs.

| Function | Input | Output | What it does |
|---|---|---|---|
| `sum()`       | A `numeric` vector | A single number | Sums the elements of a vector |
| `rev()`       | A vector         | A vector        | Reverses the elements of a vector |
| `seq()`       | Several numbers  | A vector        | Generates a sequence of numbers |
| `as.factor()` | A vector         | A vector        | Encodes a vector as `factor` type |

For example for `sum()`, if our input is

```{r}
c(3, 6, 2)
```

Our output will be the number 11.

```{r}
sum(c(3, 6, 2))
```


## Some more functions

This is a good chance to learn about some functions we haven't covered yet. If you want to know more about any of these functions, you can look up their help files.

**`abs()`**

Finds the absolute value of a number, i.e. find the distance of a number from zero (makes negative numbers positive)

```{r}
abs(-5)
abs(3.1)
```

**`round()`**

Rounds to the nearest whole number by default (the behaviour of this function when rounding off a 5 is a subtle, see the docs). If the `digits` parameter is a positive number, it will round to that number of decimal places; set to a negative number to round to powers of 10.

```{r}
round(5.233)
round(5.233, digits = 2)
round(3423.23, digits = -2)
# subtle behaviour rounding a 5 - goes to even number
round(3.5, digits = 0)
round(2.5, digits = 0)
```

**`floor()`**

Rounds, but always goes to the lowest number.

```{r}
floor(4.87)
```

**`ceiling()`**

Similar to `floor()` but always goes to the highest number.

```{r}
ceiling(3.23)
```

**`paste()`**

Joins character strings together, using whatever character separator (`sep`) you specify. Joins with a space by default.

```{r}
paste("Hello", "Goodbye")
```

```{r}
paste("Hello", "Goodbye", sep = "")
```

If you have a vector of character strings, you can join them together by specifying the `collapse` argument.

```{r}
letters = c("h", "e", "l", "l", "o")
paste(letters, collapse = "-")
```

A shortcut if you want to joining with no separator, i.e. `sep = ""`, is to use `paste0()`:

```{r}
paste0("Hello", "Goodbye")
```

**`sort()`**

Returns elements of your vector in order. This is ascending order for numbers and alphabetical order for characters, although we can change this behaviour via the `decreasing` parameter
 
```{r}
sort(c(4, 2, 5, 1))
```

```{r}
sort(c("banana", "apple", "carrot"))
```

**`nchar()`**

Finds the number of characters in a string.

```{r}
nchar("hello")
```
