---
title: "Combining events and Conditional probability"
output: html_notebook
---

```{r}
library(prob)
library(tidyverse)
```


# Calculating the complement of an event

The probability of $A^{c}$, $\textrm{prob}(A^c)$ is simply equal to $1-\textrm{prob}(A)$ since an event **has to either occur or not occur**, so 

$$\textrm{prob}(A)+\textrm{prob}(A^c) = 1$$.


```{r}
s_four_coins <- tosscoin(4) %>% 
  mutate(row = row_number())

s_four_coins
```


```{r}
s_four_coins <- s_four_coins %>% 
  mutate(num_heads = rowSums(. == "H"))

s_four_coins
```

```{r}
s_at_least_two_heads <- s_four_coins %>% 
  filter(num_heads >= 2)

prob_two_plus_heads <- nrow(s_at_least_two_heads) / nrow(s_four_coins)

prob_two_plus_heads

# get the observations where there are less than 2 heads
# s_at_least_two_heads <- s_four_coins %>% 
#   filter(num_heads < 2)

# calculate the probability
# prob_less_two <- nrow(s_less_than_two_heads) / nrow(s_four_coins)
# prob_less_two
```

# AND combinations

In probability theory, we call this the **intersection** $\cap$ of the two event spaces.
$$(A \cap B) = \{3,4\}$$

### Task - 5 mins

Given our four coin tossing example from above, what is the probability of tossing at least one head and at least one tail?

    To do this, you’ll need to add a num_tails column to s_four_coins
    Then think about filtering to satisfy the two conditions
```{r}
s_four_coins <- s_four_coins %>% 
  mutate(num_tails = rowSums(. == "T"))

s_four_coins
```

```{r}
s_four_coins %>% 
  filter(num_heads > 0 &
           num_tails > 0) 
```

```{r}
s_four_coins %>% 
  filter(num_heads > 0 &
           num_tails > 0) %>% 
  summarise(prob = n() / nrow(s_four_coins))
```

```{r}
# gives the same as using &
s_four_coins %>% 
  filter(num_heads > 0,
           num_tails > 0)
```

# OR combinations

In probability theory, this involves taking the **union** of the two event spaces:

$$(A \cup B) = \{2,4,5,6\}$$

```{r}
# create our sample space: two die rolls
s_two_dice <- rolldie(2)

# now filter down to cases where X1 is even or X2 is 5
s_two_dice %>% 
  filter(X1 %% 2 == 0 | X2 == 5) %>% 
  summarise(prob = n() / nrow(s_two_dice))
```

### Optional Task - 10 mins

Let's check the addition law for ourselves. Calculate probabilities for each event below:

* $\textrm{first_die_even}$
* $\textrm{second_die_5}$
* $\textrm{first_die_even AND second_die_5}$

Then use these probabilities in the expression from above

$$\textrm{prob}(A \cup B) = \textrm{prob}(A) + \textrm{prob}(B) - \textrm{prob}(A \cap B)$$

and check you get the same probability for $\textrm{prob}(\textrm{first_die_even OR second_die_5})$ as we obtained from the last block of code.

<details>
<summary>**Hint**</summary>
$$
\begin{multline}
\textrm{prob}(\textrm{first_die_even OR second_die_5}) = \\ 
\textrm{prob}(\textrm{first_die_even}) + \textrm{prob}(\textrm{second_die_5}) \\
- \textrm{prob}(\textrm{first_die_even AND second_die_5}) 
\end{multline}
$$

```{r}
first_die_even <- s_two_dice %>%
  filter(X1 %% 2 == 0) %>% 
  summarise(prob = n() / nrow(s_two_dice)) %>% 
  pull()

second_die_5 <- s_two_dice %>% 
  filter(X2 == 5) %>% 
  summarise(prob = n() / nrow(s_two_dice)) %>% 
  pull()

both_con <- s_two_dice %>% 
  filter(X1 %% 2 == 0 & X2 == 5) %>% 
  summarise(prob = n() / nrow(s_two_dice)) %>% 
  pull()

pr_union <- first_die_even + second_die_5 - both_con 

# pull() takes the tibble df and makes it a value to use in the sum above
```


# Conditional probability

For Conditional probability | is **given that** not **or**

In terms of probability theory, the conditional probability is

$$\textrm{prob}(A \mid B) = \frac{\textrm{prob}(A \cap B)}{\textrm{prob}(B)}$$

```{r}
cb <- read_csv("data/colour_blindness.csv")
```

```{r}
head(cb)
```

```{r}
library(janitor)
```

```{r}
table <- cb %>% 
  tabyl(bio_sex, cblind)

table

# takes the 
```

```{r}
table <- table %>% 
  adorn_totals(where = c("row", "col"))

table
```

```{r}
table <- table %>%
  adorn_percentages(denominator = "all") %>%
  adorn_title()

table
```

```{r}
prob_female <- cb %>%
  summarise(prob = sum(bio_sex == "female") / n())
prob_female

# shows the same prob of being Female
```

### Task - 5 mins

What are the marginal probabilities of:

    being male
    being colour blind
    not being colour blind?

Get these probabilities from the contingency table, and also calculate them using pipes. Save each probability in a sensibly named variable, we’ll need it later!

```{r}
prob_male <- cb %>%
  summarise(prob = sum(bio_sex == "male") / n())
prob_male
```

```{r}
prob_cb <- cb %>%
  summarise(prob = sum(cblind) / n())
prob_cb
```

```{r}
prob_not_cb <- cb %>% 
  summarise(prob = sum(!cblind) / n())
prob_not_cb
```

# Joint probabilities

```{r}
prob_female_cb <- cb %>% 
  filter(bio_sex == "female" & cblind == TRUE) %>% 
  summarise(prob = n() / nrow(cb))

prob_female_cb
```

# Dependent and independent events

```{r}
prob_female * prob_cb == prob_female_cb
```

```{r}
prob_female * prob_cb == 0.01744	

prob_female_cb == 0.0016
```

This is independent

# Conditional probability


```{r}
prob_cb_given_female = cb %>%
  filter(bio_sex == "female") %>%
  summarise(prob = sum(cblind) / n())

prob_cb_given_female
```

###Task - 5 mins

Over to you again! Calculate $\textrm{prob}(\textrm{colour blind | male})$ first by using the formula

$$\textrm{prob}(\textrm{colour blind | male}) = \frac{\textrm{prob}(\textrm{colour blind} \cap \textrm{male})}{\textrm{prob}(\textrm{male})}$$
```{r}
prob_cb_given_male = cb %>%
  filter(bio_sex == "male") %>%
  summarise(prob = sum(cblind) / n())

prob_cb_given_male
```




