---
title: "Core definitions and Counting rules"
output: html_notebook
---

# Core definitions

### Task - 5 mins Two dice are rolled, leading to a sample space:

Stwo-dice={(1,1),(1,2),(1,3),(1,4)...,(6,4),(6,5),(6,6)}
.


Write out the set of outcomes corresponding to the event that the total of the two dice is 7.

Hint: You want to write out all possible combinations of dice rolls that would add up to 7

(1,6), (2,5), (3,4), (4,3), (5,2), (6,1)


```{r}
library(prob)
library(tidyverse)
```


```{r}
# sample space for roll of two dice
s_two_dice <- rolldie(2)
s_two_dice
```

```{r}
# event space for roll of a total of seven
s_sum_equals_seven <- s_two_dice %>%
  filter(X1 + X2 == 7)

# how large is the event space?
s_sum_equals_seven %>%
  summarise(size_of_event_space = n())

s_sum_equals_seven
```

random variables X1 and X2 use caps, wacth out when cleaning data

```{r}
s_two_dice <- s_two_dice %>%
  mutate(total = X1 + X2)
s_two_dice
```

```{r}
s_two_dice %>% 
  count(total)
```


### Task - 5 mins

Add your own random variable to s_two_dice, and then use your new random variable to define an event space. Some suggestions:

    random variable: first_die_even, event: first_die_even == TRUE
    random variable: die2_minus_die1, event: die2_minus_die1 == -2
    random variable: differ_by_two, event: differ_by_two == FALSE


```{r}
s_two_dice <- s_two_dice %>%
  mutate(first_die_even = X1 %% 2 == 0) %>%
  mutate(die2_minus_die1 = )


s_two_dice
```

# Counting rules

```{r}
# get all possible combos of 2 die rolls
s <- rolldie(2)

# find when the two separate die rolls total 7
total_seven <- s %>%
  filter(X1 + X2 == 7)

total_seven
```

```{r}
# number of outcomes / total number of options
prob_total_seven <- nrow(total_seven) / nrow(s)
prob_total_seven
```

### What is the probability of getting heads on a coin toss and a 6 on a die throw?

```{r}
# set up one roll of a dice
s_die <- rolldie(1)

# set up one toin coss
s_coin <- tosscoin(1)

# merge them together
s_total <- merge(s_coin, s_die)

#how many outcomes for our question?
head_six <- s_total %>%
  dplyr::filter(toss1 == "H", X1 == 6)

# calculate the probability
prob_head_six <- nrow(head_six) / nrow(s_total)
prob_head_six
```

```{r}
?intersect()
# Finds the comming columns 

intersect(names(s_coin), names(s_total))
# "toss1" is comming between the both
```


## Permutations

The permutations counting rule formula is as follows:
$$^NP_r = \frac{N!}{(N-r)!}$$

We could say this in plain language as:

‘The number of permutations of N things taken r at a time’

```{r}
library(gtools)
```

```{r}
# create vector of choices contains ten employees, n = 10. 
employees <- c("Alice", "Bob", "Carole", "David", "Elaine", 
               "Fred", "Gabra", "Hua", "Inigo", "Hugo")

# Choose three at a time, r = 3
# Don't replace an employee once chosen: repeats.allowed = FALSE
seniors <- as_tibble(permutations(n = 10, 
                                  r = 3, 
                                  v = employees, 
                                  repeats.allowed = FALSE))

# n	Size of the source vector
# 
# r	Size of the target vectors
# 
# v	Source vector. Defaults to 1:n

# repeats.allowed = is defaulted to FALSE

# rename the variables in seniors to something better
seniors <- seniors %>%
  rename(
    CEO = V1,
    CTO = V2,
    COO = V3
  )

# let's see what the seniors matrix looks like
seniors
```

### What is the probability of Alice being selected as CEO and Bob being selected as CTO, if the selection process is random?
```{r}
# first get the event space
alice_CEO_bob_CTO <- seniors %>%
  filter(CEO == "Alice", CTO == "Bob")
alice_CEO_bob_CTO
```

```{r}
# then calculate the prob
prob <- nrow(alice_CEO_bob_CTO) / nrow(seniors)
prob

# makes it 8/270 which is 1/90 odds
```


## Combinations

We can find the total number of combinations using the following formula:

$$^NC_r = \binom{N}{r} = \frac{N!}{r! \times(N-r)!}$$
"N choose r" is how some stats people say Combinations, so the function is choose()

```{r}
# use the choose() function instead, which just computes the number of choices from 50 numbers, drawn in lots of 6
num_draws <- choose(50, 6)
num_draws
```

### Task - 5 mins

Use the combinations() function in gtools for this task (don’t try to do it manually). I have five flavours of ice cream to choose from:

flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio") 

and a bowl that can hold three scoops. How many different variations in bowls of ice cream can I have assuming:

    Each scoop must be a different flavour
    I can repeat flavours if I wish?

```{r}
flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio") 
```

```{r}
as_tibble(combinations(n = 5,
                       r = 3,
                       v = flavours,
                       repeats.allowed = FALSE))

```

```{r}
as_tibble(combinations(n = 5,
                       r = 3,
                       v = flavours,
                       repeats.allowed = TRUE))
```


