---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)
library(fastGraph)
```

### Q1
Load and explore the data held in heart.csv. A data dictionary is also available.
```{r}
heart <- read_csv("data/heart.csv")
```

```{r}
glimpse(heart)
```

### Q2
Plot a histogram and/or boxplot for thalach (maximum heart rate achieved in stress testing). How normal is the distribution?
```{r}
heart %>% 
  ggplot() +
  aes(x = thalach) +
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7) 
  
```

### Q3
Also plot a barplot of exang (whether the patient displayed exercise induced angina)

```{r}
heart %>% 
  ggplot() +
  aes(x = exang) +
  geom_bar(col = "white", fill = "steel blue", alpha = 0.7)
```

### Q4
Now calculate population parameters: what is the mean thalach in the population. What proportion of patients have exercise induced angina (exang)?

```{r}
pop_param <- heart %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = sum(exang == 1) / n()
  )

pop_param
```

### Q5
Now take a single sample of 50-patients from the population and calculate point estimates for mean `thalach` and proportion of `exang` from this sample (let’s call these `mean_thalach` and prop_exang going forward).
[Hint - use `rep_sample_n()` from the `infer` package to do this]
```{r}
sample_50 <- heart %>% 
  rep_sample_n(size = 50, reps= 1)

summ_sample_50 <- sample_50 %>% 
  ungroup() %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = sum(exang == 1) / n()
  )

summ_sample_50
```



### Q6
What are the sampling errors of these point estimates?
[Hint - remember a sampling error is defined as the difference between a point estimate drawn from a sample, and a population parameter]
```{r}
sampling_errors <- summ_sample_50 - pop_param

sampling_errors
```

### Q7
Now we’re going to repeatedly sample 50-patients from the population and build up sampling distributions for the mean_thalach and prop_exang point estimates. Take 5000 samples and store the sampling distributions in a variable rep_sample_50.
[Hint - again, use rep_sample_n() to do this.]
```{r}
resample_50 <- heart %>% 
  rep_sample_n(size = 50, reps= 5000)

rep_sample_50 <- resample_50 %>% 
  #ungroup() %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = mean(exang)
  )

head(rep_sample_50)
```

### Q8
Visualise the sampling distributions of `mean_thalach` and `prop_exang` as histograms. How normal do the sampling distributions look?
```{r}
mean_thalach_plot <-  rep_sample_50 %>% 
  ggplot(aes(x = mean_thalach)) +
  geom_histogram(col = "black", fill = "red", alpha = 0.7) +
  labs(x = "mean thalach")

mean_thalach_plot
```

```{r}
prop_exang_plot <-  rep_sample_50 %>% 
  ggplot(aes(x = prop_exang)) +
  geom_histogram(col = "black", fill = "red", alpha = 0.7) +
  labs(x = "proportion of exang")

prop_exang_plot
```

### Q9 
Obtain standard errors of the sampling distribution of mean_thalach and prop_exang directly as the standard deviations of the sampling distributions held in rep_sample_50
```{r}
stand_error <- rep_sample_50 %>% 
  summarise(se_mean_thalach = sd(mean_thalach),
            se_prop_exang = sd(prop_exang)
  )

stand_error
```

### Q10
Repeat steps 7. and 9. (omitting step 8.) once more with samples of size 100-patients, and again with samples of size 200-patients. How do the standard errors of mean_thalach and prop_exang change as sample size increases? Does the uncertainty in the calculated statistics increase or decrease as sample size increases?
[Hint - it’s OK to copy, paste and alter code to do this!]
```{r}
resample_100 <- heart %>% 
  rep_sample_n(size = 100, reps= 5000)

rep_sample_100 <- resample_100 %>% 
  #ungroup() %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = mean(exang)
  )

rep_sample_100

stand_error_100 <- rep_sample_100 %>% 
  summarise(se_mean_thalach = sd(mean_thalach),
            se_prop_exang = sd(prop_exang)
  )

stand_error_100
```
```{r}
resample_200 <- heart %>% 
  rep_sample_n(size = 200, reps= 5000)

rep_sample_200 <- resample_200 %>% 
  #ungroup() %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = mean(exang)
  )

rep_sample_200

stand_error_200 <- rep_sample_200 %>% 
  summarise(se_mean_thalach = sd(mean_thalach),
            se_prop_exang = sd(prop_exang)
  )

stand_error_200
```


